### 2024-04-14 R0_journal/main.tex
fixed animation and open loop codesign works better

\documentclass[smallextended]{./springer/svjour3}
\usepackage{graphicx} % This lets you include figures
\graphicspath{ {./figures/} }
\usepackage[rightcaption]{sidecap}
% \usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{imakeidx}
\usepackage{comment}
\usepackage{commath}
\usepackage[titletoc]{appendix}
\usepackage{graphicx}
\usepackage{resizegather}
\usepackage[english]{babel}
% \usepackage{subcaption}
\usepackage{tabu}
\usepackage{booktabs}
\usepackage{xfrac}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsmath}
% \usepackage{amsthm}
\usepackage{commath}
\usepackage{graphicx,bm}
\usepackage{verbatim}
% \usepackage{caption}
\usepackage{lscape}
\usepackage{relsize}
\usepackage{enumitem}
\usepackage{textcomp}
\usepackage{breqn}
\usepackage{makecell}
\usepackage{longtable,tabularx}
\usepackage{multirow}
\usepackage{doi}
\usepackage{fancyhdr}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{setspace}
\usepackage{footnote}
\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue}}
\usepackage[numbers]{natbib}
\usepackage{mathtools}
\usepackage[disable]{todonotes}
\usepackage[framemethod=tikz]{mdframed}
\usepackage{booktabs,xcolor,siunitx}
\usepackage{soul}
% \usepackage{cleveref}
\usepackage[small, compact]{titlesec}
\usepackage{xcolor}
\usepackage{appendix}
% \usepackage{geometry}
% \geometry{
% a4paper,
% total={170mm,257mm},
% left=20mm,
% top=20mm,
% }

% \newtheorem{theorem}{Theorem}[section]
% \newtheorem{lemma}[theorem]{Lemma}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\degree}{\ensuremath{^\circ\,}}
\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
\newcommand{\f}[2]{\frac{#1}{#2}}
\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\tr}[1]{\mathrm{Tr}\left({#1}\right)}
\newcommand{\mbg}[1]{\boldsymbol{\mathbf{#1}}}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}
\renewcommand{\d}{\mathop{}\!\mathrm{d}} % total derivative
\newcommand{\p}{\partial}

\title{Research project title}
\author{Max Howell}
\institute{University Of Tennessee Knoxville$^*$ ($^*$corresponding author),
          \email{mhowel30@vols.utk.edu} \\
        \\
          \at MABE, University of Tennessee, Knoxville,
          \at Nathan W. Dougherty Engineering Building, 1512 Middle Dr, Knoxville, TN 37916\\
}

\date{\today}

\begin{document}
\maketitle{}

\begin{abstract}
Description of the project.
\end{abstract}

\section{Problem setup}

\end{document}

### 2024-04-14 code/compass-gait/animate.py

@@ -4,16 +4,12 @@ import matplotlib.patches as patches
 from matplotlib import animation
 
 
-def animate_compass(x1, x2, a, b, phi, name='compass.gif', interval = 20, saveFig=False, gif_fps=20):
+def animate_compass(x1, x2, a, b, phi, name='compass.gif', interval = 20, saveFig=False, gif_fps=20, iter=1, num_iter_points=90):
     #x1: back leg
     #x2: front leg
     #phi: angle of slope
 
     l = a + b
-
-    extend = 500 // interval
-    x1 = np.concatenate((x1, np.ones(extend) * x1[-1]))
-    x2 = np.concatenate((x2, np.ones(extend) * x2[-1]))
 
     # path of stance leg tip (hip)
     x_hip = -l*np.sin(x2)
@@ -32,6 +28,44 @@ def animate_compass(x1, x2, a, b, phi, name='compass.gif', interval = 20, saveFi
     #x_lim = [min(min(x), min(x_pole)) - cart_width / 2 - 0.1, max(max(x), max(x_pole)) + cart_width / 2 + 0.1]
     #ylim = [cart_height / 2 - l - 0.05, cart_height / 2 + l + 0.05]
 
+    extend = 500 // interval # extend all paramaters 
+
+    if (iter>1): # make animation walk forward
+        base_x = np.zeros(num_iter_points*iter)
+        base_y = np.zeros(num_iter_points*iter)
+        base_x[0:num_iter_points] = 0
+        base_y[0:num_iter_points] = 0
+        for i in range(iter):
+            index_prev = i*num_iter_points # end index of cycle i
+            index_next = (i+1)*num_iter_points # start index of cycle i
+            #if i>0:
+            #    base_x[index_prev:index_next] = -x_swing[index_prev] # if cycle is not the first cycle, make new x and y base for stance leg
+            #    base_y[index_prev:index_next] = -y_swing[index_prev]
+
+            x_hip[index_prev:index_next] = -l*np.sin(x2[index_prev:index_next]) + base_x[index_prev:index_next]
+            y_hip[index_prev:index_next] = l*np.cos(x2[index_prev:index_next]) + base_y[index_prev:index_next]
+
+            x_hip2swing[index_prev:index_next] = l*np.sin(x1[index_prev:index_next])
[…]

### 2024-04-14 code/compass-gait/compass_torque.py

@@ -14,7 +14,7 @@ class system_active(om.Group):
 
         input_names = ['a', 'b', 'x1', 'x2', 'x3', 'x4', 'mh', 'm', 'tau']
         self.add_subsystem('lockedknee', dynamics(num_nodes=nn, ), promotes_inputs=input_names, promotes_outputs=['*'])
-        self.add_subsystem('cost', CostFunc(num_nodes=nn, states_ref=self.options['states_ref'] ), promotes_inputs=['time', 'mh'], promotes_outputs=['*'])
+        self.add_subsystem('cost', CostFunc(num_nodes=nn, states_ref=self.options['states_ref'] ), promotes_inputs=['tau', 'time'], promotes_outputs=['*'])
 
 
 
@@ -202,35 +202,35 @@ class CostFunc(om.ExplicitComponent):
 
     def setup(self):
         nn = self.options['num_nodes']
-        #self.add_input('tau', shape=(nn,), units='N*m', desc='input torque')
-        self.add_input('mh', shape=(1,), units='kg', desc='hip mass')
+        self.add_input('tau', shape=(nn,), units='N*m', desc='input torque')
+        #self.add_input('mh', shape=(1,), units='kg', desc='hip mass')
         self.add_input('time', shape=(nn), units='s', desc='time')
 
-        self.add_output('costval', shape=(nn,), desc='cost value')
+        self.add_output('costrate', shape=(nn,), desc='cost value')
         
-        self.declare_partials(of=['costval'], wrt=['time'], method='exact', rows=np.arange(nn), cols=np.arange(nn))
+        self.declare_partials(of=['costrate'], wrt=['tau', 'time'], method='exact', rows=np.arange(nn), cols=np.arange(nn))
         #self.declare_coloring(wrt=['m_H','m_t','m_s', 'tau',], method='cs', show_summary=False)
         #self.set_check_partial_options(wrt=['m_H','m_t','m_s', 'tau',], method='fd', step=1e-6)
 
-        #self.declare_partials(of=['costval'], wrt=['time'], method='exact')
+        #self.declare_partials(of=['costval'], wrt=['mh'], method='exact')
 
     def compute(self, inputs, outputs,):
-        #tau = inputs['tau']
+        tau = inputs['tau']
         #mh = inputs['mh']
         time = inputs['time']
 
-        outputs['costval'] = time #- (mh/150)
+        outputs['costrate'] = ((tau**2)/40) + (time/0.5)
         
 
[…]

### 2024-04-14 code/compass-gait/run_withsim.py

@@ -8,6 +8,10 @@ from dymos.examples.plotting import plot_results
 import numpy as np
 from numpy.linalg import inv
 
+"""
+OPEN LOOP CONTROL OVER LIMIT CYCLE (CO-DESIGN FOR FIRST CYCLE ONLY) WITH OBJ: MINIMIZE TIME AND TAU WITH TRAJECTORY GENERATION
+"""
+
 # defining paramters of the legged walker
 a = 0.5
 b = 0.5
@@ -114,7 +118,7 @@ def traj_gen():
                   ('traj.initphase.timeseries.time','traj.initphase.timeseries.states:x4','time','q2_dot'),
                   ('traj.initphase.timeseries.time', 'traj.initphase.timeseries.parameters:mh', 'time', 'hip mass')],
                   title='Time History',p_sol=p,p_sim=sim_sol)
-    plt.savefig('compass_passive_opt=false.pdf', bbox_inches='tight')
+    plt.savefig('compass_trajgen', bbox_inches='tight')
 
     # add end conditions to be used with control
     x1_end = p.get_val('traj.initphase.states:x1')[-1]
@@ -145,7 +149,7 @@ def traj_gen():
     ax.set_ylabel('angular velocity')
     #ax.legend()
 
-    plt.savefig('limitcycle_opt=false.pdf')
+    plt.savefig('limitcycle_trajgen')
 
     ## plot motion
 
@@ -157,7 +161,7 @@ def traj_gen():
 
     #plot animation
     from animate import animate_compass
-    animate_compass(x1arr.reshape(num_points), x2arr.reshape(num_points), a, b, phi, name='opt_false.gif', saveFig=True)
+    animate_compass(x1arr.reshape(num_points), x2arr.reshape(num_points), a, b, phi, name='trajgen.gif', saveFig=True, gif_fps=20)
 
     
     return newstates_final
@@ -166,7 +170,7 @@ def traj_gen():
 def traj_opt(a=0.5, b=0.5, mh=10, m=5, states_final=[0, 0, 0, 0], states_init=[0, 0, 0, 0], co_design=True):
 
     #defining paramters of the legged walker
-    duration_lockphase = 1
+    duration_lockphase = 0.5
     a = a
     b = b
     mh = mh
@@ -181,6 +185,7 @@ def traj_opt(a=0.5, b=0.5, mh=10, m=5, states_final=[0, 0, 0, 0], states_init=[0
     optimized vars to planned trajectory
     """
[…]

### 2024-05-07 R0_journal/main.tex
paper

@@
 \usepackage{tabularx}
-\usepackage{amssymb}
+% \usepackage{amssymb}
 \usepackage{amsmath}
@@
 \usepackage{mathtools}
-\usepackage[disable]{todonotes}
+%\usepackage[disable]{todonotes}
 \usepackage[framemethod=tikz]{mdframed}
@@
 
+%\usepackage[section]{placeins}
+
 \title{Research project title}
@@
 \begin{document}
+\bibliographystyle{plain}
+\bibliography{references}
+
 \maketitle{}
@@
 \begin{abstract}
-Description of the project.
+
+Control Co-Design (CCD) is a method used to optimally design mechanical systems which involves desinging the systems parameters and controling the system at the same time. This method 
+had potential to be less time consuming and more efficent than traditional methods used to design systems where an iterative process is used. In this paper, CCD is applied to a 
+bipedal walking robot known as the compass gait in order to optimize the masses of the upper body and legs and the length of the legs in order to reduce the torque needed to control 
+the system and reduce the time per cycle when walking. 
+A trajectory generation method was also used in this paper to ensure stability of the compass gait system, as the passive compass gait has a very small range of initial conditions for 
+which the system is stable for.
+During the optimization process the cost of control for the compass gait was reduced by 10.83$\%$
+when compared to a compass gait using only open loop control. The torque needed to control the compass gait was also reduced by 78.2$\%$ and the time per cycle was reduced by 10.6$\%$.
+It was also shown that even a simple optimization problem of the passive compass gait can increase the range of stability for the system. When a passive compass gait was optimized to maximize the hip 
+mass, it was shown that not only can the hip mass be optimized by chaning the system parameters, but when simulating the optimized compass gait with 
+the same initial conditions which cause a unstable passive system, the optimized system was stable.
+
 \end{abstract}
 
-\section{Problem setup}
+\section{Introduction}
+% introduce control-codesign
+Control Co-Design (CCD) is a design method used to optimally design a system's parameters and control the system at the same time. Traditionally in engineering,
+when a system is designed an iterative process is used to determine the best parameters for the system; however, this can be time consuming and inefficient. 
+It has been shown in \cite{} that the cost of control can be decreased by atleast xx\% in systems such as a cartpole pendulum and a quadrotor by using CCD. CCD applications range
+from design optimizatoin of (examples of CCD being used with citations)
+
+%introduce dymos a a way to use CCD for walking bipdel system
+CCD can be used in junciton with optimal control algorithms with the objective of minimizing or maximizing a desired cost function. One 
+excellent resource used in this paper to perform both CCD and optimal control in a bipedal walking robot is Dymos \cite{Falck2021}. The Dymos multidisciplinary optimal control library
+provides a platform to build and control systems with a variety of different control algorithms.
+
+%
+In this paper, CCD is used to design the system parameters of an underactuated bipedal walking robot known as the compass gait. The limit cylce
+of the compass gait model aswell as the effect of changing system parameters has been thoroughly studied 
+in \cite{Goswami} and \cite{goswami:inria-00073701}. Control of the compass gait robot as also been acheived using passive methods such as potential energy shaping
+\cite{Spong} and active methods such as MPC Control \cite{Kamath2009}; however, CCD has not been applied to the compass gait model.
+In this paper CCD is applied to the compass gait model using open loop control and trajectory optimization with Dymos.
+
+\section{Introduction to the Compass Gait}
+The compass gait model is an underactuated two legged robot capable of passively walking downhill under the right initial conditions.
+Shown in Figure \ref{fig:compassgaitmodel} is a schematic showing the compass gait and its respective system parameters. The two leg masses are modeled as
+point masses with leg lengths a and b below and above the point mass making a total leg length l. The upper body of the robot is approximated as a point mass
+applied at the hip, $m_H$. The state variables of the system are the swing leg angle (the back leg) $\theta_{sw}$ and the stance leg (the front leg) angle
+$\theta_{st}$. A torque, $\tau$ is applied at the hip to each leg in the opposite direction. The angle of the ramp which the compass gait walks on is $\gamma$.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=10cm]{./figures/compassgaitmodel.png}
+\caption{Schematic of the compass gait model. Angles $\theta_{sw}$ and $\theta_{st}$ are shown as the swing leg angle and stance leg angle, respectively.}
+\label{fig:compassgaitmodel}
+\end{figure}
+
+% assumptions for model
+In order to use the compass gait model a few assumptions must be made about the system. First, since the compass gait does not have the ability to bend at the hip, it must 
+be assumed that during the swing, the swing leg retracts so that it does not collide with the ground during the swing phase. 
+It is also assumed that a perfectly inelastic collision without sliding
+occurs the instant the swing leg hits the ground. The compass gait is an extremely simple model, however it should be noted that physical models of the compass gait
+have been built in \cite{article}.
+
+% intro of eqs
+In order to create a state space realization of the compass gait model a state matrix must be defined. For simplicity the swing and stance angles are renamed to
+$x_1$ and $x_2$, respectively, with $x_3$ and $x_4$ being the anguar velocities of the swing and stance legs. The state matrix $\mb{X}$ is shown in equation
+\ref{eq:statematrix}.
+
+\begin{equation}
+\label{eq:statematrix}
+\mb{X} = 
+\begin{bmatrix}
+\\x_1\\
+\\x_2\\
+\\x_3\\
+\\x_4\\
+\end{bmatrix}, \quad
+\begin{bmatrix}
+\\x_3\\
+\\x_4\\
+\end{bmatrix}
+ = 
+\begin{bmatrix}
+\\\dot{x_1}\\
+\\\dot{x_2}\\
+\end{bmatrix}
+\end{equation}
+
+The dynamic equations of the swing stage of the compass gait, from \cite{french confrence}, can be written in standard planar manupular dynamics form. 
+This is shown in \ref{eq:swingeqn}, where $\mb{M}$, $\mb{N}$, and $\mb{G}$ are matrices shown in the appendix, and $\mb{U}$
+ is the vector used to apply the torque to each leg, shown in
+\ref{eq:U}.
+
+\begin{equation}
+\label{eq:swingeqn}
+\mb{M(X)}\mb{\ddot{X}} + \mb{N(X, \dot{X})}\mb{\dot{X}} + \mb{G(X)} = \tau\mb{U}
+\end{equation}
+
+\begin{equation}
+\label{eq:U}
+\mb{U} = 
+\begin{bmatrix}
+\\1\\
+\\-1\\
+\\0\\
+\\0\\
+\end{bmatrix}
+\end{equation}
+
+Since heel strike (when the swing leg collides with the ground) is modeled as a perfectly inelastic 
+collision, there is an instantaneous angular velocity change when heel stike occurs. Also, since the swing leg becomes the stance leg
+after heel strike, the angles $x_1$ and $x_2$ must be switched in the mathematical model after each walking cycle. The transition matrices used to switch $x_1$ and $x_2$ as well 
+as calculate the updated angular velocities after heel strike are shown in \ref{eq:trans1} and \ref{eq:trans2}, where $x^+$ represents the state
+variable after heel strike and $x^-$ represents the state variable before impact, and $\mb{H}$ is the angular velocity transition matrix shown in the appendix.
+
+\begin{equation}
+\label{eq:trans1}
+\begin{bmatrix}
+\\x_1^+\\
+\\x_2^+\\
+\end{bmatrix}
+ = 
+\begin{bmatrix}
+0 & 1\\
+1 & 0\\
+\end{bmatrix}
+\begin{bmatrix}
+\\x_1^-\\
+\\x_2^-\\
+\end{bmatrix}
+\end{equation}
+
+\begin{equation}
+\label{eq:trans2}
+\begin{bmatrix}
+\\x_3^+\\
+\\x_4^+\\
+\end{bmatrix}
+ = 
+\mb{H}
+\begin{bmatrix}
+\\x_3^-\\
+\\x_4^-\\
+\end{bmatrix}
+\end{equation}
+
+\section{Natural Limit Cycles of the Passive Compass Gait}
+
+It is important to define stability in the context of the compass gait for this paper. One possible meaning of stability in this context is that, while walking, 
+the compass gait procedes forwards in a controlled manner and takes the same "path" for each walking cycle. Or perhaps stability could be mathematically defined
+using Poincaré maps. While both of these definitions of stability are valid, stability in this paper simply refers to the ability of the compass gait system to 
+continue walking forwards for the defined number of cycles.
+
+When simulating the passive compass gait system, it can be observed that the stability of the compass gait is heavily dependent on the
+given initial conditions and system parameters. It can also be observed that the basin of stability is extremely small and it is almost impossible
+to "guess" the correct IC's and system parameters to create a stable walking system. This is one of the main issues of the compass gait.
+
+When using stable initial conditions, the compass gait can be simulated over one or multiple walking cycles. Shown in Figure \ref{fig:onecycle}
+is an example of a stable limit cycle for the compass gait over one walking cycle. As can be seen, the swing leg makes a large arc during the swing phase of the 
+compass gait, while the stance leg pivots in place. The state variables during this walking cycle are also shown over time in Figure \ref{fig:states_onecycle}.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=9cm]{./figures/onecycle.pdf}
+\caption{One walking limit cycle of the compass gait. The swing leg makes a large swing (shown in blue, a large cicle) around while the stance leg only pivots (shown in orange.)}
+\label{fig:onecycle}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=10cm]{./figures/states_onecycle.pdf}
+\caption{States of the compas gait over time for one walking cycle.}
+\label{fig:states_onecycle}
+\end{figure}
+
+Figures \ref{fig:onecycle} and \ref{fig:states_onecycle} show the compass gait when stable initial conditions are used; however, most initial conditions
+given to the compass gait system result in an unstable system. An example of a unstable gait is shown in Figure \ref{fig:unstable_onecycle}, where the initial angular velocities are 
+not sufficent to overcome gravity, and the compass gait cannot start a cycle and will eventually fall down. It can be seen in this figure, that the swing leg initial angle 
+is approximataly equal to the final angle of the swing leg, and the stance leg starts going forward, but runs out of momentum before falling backwards.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=10cm]{./figures/unstable_onecycle.pdf}
+\caption{An example of an unstable limit cycle. In this case, the initial angular velocities make it impossible for the compass gait to start a walking cycle.}
+\label{fig:unstable_onecycle}
+\end{figure}
+
+\section{Passive Optimization with Co-Design}
+
+In order to acheive co-design with a simple passive compass gait system, an optimization problem was proposed.
+This optimization problem, outlined in Figure \ref{fig:optsetup}, involved maximizing the point mass at the hip ($m_H$) in order to simulate the
+compass gait robot "carrying" a heavy object. In order to acheive this goal, the lengths below and above the leg point mass, a and b, were optimized.
+The leg point mass ($m$) also changed based on the total length of the legs based on the given density of the legs.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=6cm]{./figures/opt1prob.png}
+\caption{Outline of the simple passive optimization problem. The goal is to incrase the range of stable initial conditions while maximizing the applied hip mass.}
+\label{fig:optsetup}
+\end{figure}
+
+The goal of this optimization problem was to increase the range of initial conditions that caused the compass gait to complete one stable cycle
+while maximizing the mass at the hip. In order to acheive this goal a trajectory generation method was used, where the compass gait model was
+first simulated using known stable initial conditions and system parameters and the final position states ($x_1$ and $x_2$) were recorded and passed to the optimization
+model. The final position states of the optimization model were constrained to be equal to the recorded final position states found in the trajectory generation
+simulation. This gave the the compass gait optimization problem a goal to acheive in order to keep the system stable. Figure \ref{fig:optlogicdiagram} shows a diagram
+outlining this process.
+
+\begin{figure}[h]
+\setlength{\unitlength}{0.14in} % selecting unit length
+\centering % used for centering Figure
+\begin{picture}(32,15) % picture environment with the size (dimensions)
+% 32 length units wide, and 15 units high.
+\put(0,5.5){\vector(1,0){5}}
+\put(5,3.25){\framebox(10,5){Trajectory Generation}}
+\put(20.2,3.25){\framebox(10,5){Optimization Problem}}
+\put(15,5.5){\vector(1,0){5.2}}
+\put(30.2,5.5){\vector(1,0){12}}
+\put(25.2, 12){\vector(0,-1){3.7}}
+\put(0,6){Stable IC's}
+\put(15.2, 6){End States}
+\put(15.3, 4.5){$x_1$ and $x_2$}
+\put(30.4, 6){Stable, optimized system}
+\put(23.3, 12.2){New IC's}
+\end{picture}
+\caption{A trajectory generation method was used where a system with stable initial conditions found end states ($x_1$ and $x_2$) to be used for the optimization problem.} % title of the Figure
+\label{fig:optlogicdiagram} % label to refer figure in text
+\end{figure}
+
+While it was not possible to make all initial conditions stable without a controller, the range of initial conditions was
+increased. Shown in Figure \ref{fig:nonoptlimcycle} is the limit cycle of the compass gait without optimization with the initial conditions and system parameters 
+shown in Table \ref{tab:nonoptparam}. As can be seen, the compass gait system is not able to complete a full cycle and can be said to be unstable.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=10cm]{./figures/nonopt_limitcycle.pdf}
+\caption{Limit cycle of the non optimized compass gait over one cycle with the same initial conditions as the optimized system. It can be seen that the system does not 
+complete a full cycle.}
+\label{fig:nonoptlimcycle}
+\end{figure}
+
+\begin{table}[h]
+\centering
+\caption{Initial Conditions and System Parameters of the Non-Optimized Compass Gait Shown in \ref{fig:nonoptlimcycle}}
+\begin{tabular}{lr}
+\toprule
+Parameter & Value \\
+\midrule
+$m (kg)$ & 5 \\
+$m_H (kg)$ & 10 \\
+$a (m)$ & 0.5 \\
+$b (m)$ & 0.5 \\
+$\gamma (slope, rad)$ & 0.05 \\
+$x_1 (rad)$ & -0.23 \\
+$x_2 (rad)$ & 0.4 \\
+$x_3 (rad/s)$ & -0.24 \\
+$x_4 (rad/s)$ & -1.1 \\
+\bottomrule
+\end{tabular}
+\label{tab:nonoptparam}
+\end{table}
+
+The limit cycle of the non-optimized system can be compared to that of the optimized system in order to see the benefits of the optimization.
+Shown in Figure \ref{fig:passiveopt} is the limit cycle of the optimized compass gait system with the same initial conditions as the non-optimized compass gait.
+It can be seen that after optimization, the system completes a full stable walking cycle.
+The optimized system parameters of the new system are shown in Table \ref{tab:optparam}. Not only can it be seen that the compass gait system gained stability from the 
+optimization, but it also succesfully maximized the mass at the hip.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=10cm]{./figures/passiveopt_limitcycle.png}
+\caption{The limit cycle of the optimized compass gait. It can be seen that the system is stable for the first walking cycle.}
+\label{fig:passiveopt}
+\end{figure}
+
+\begin{table}[h]
+\centering
+\caption{Initial Conditions and System Parameters of the Non-Optimized Compass Gait Shown in \ref{fig:nonoptlimcycle}}
+\begin{tabular}{lr}
+\toprule
+Parameter & Value \\
+\midrule
+$m (kg)$ & 8.96 \\
+$m_H (kg)$ & 80 \\
+$a (m)$ & 0.83 \\
+$b (m)$ & 0.957 \\
+\bottomrule
+\end{tabular}
+\label{tab:optparam}
+\end{table}
+
+While this optimization problem was succesful, it has its limitations. First, as previously mentioned, without a using an active controller
+it is impossible to guarentee that all initial conditions for the compass gait will result in a stable walking limit cycle. Also, it is desired that the optimized
+compass gait system complete multiple walking cycles instead of just one. In order to fix these two limitation, active CCD can be used to optimize the problem 
+and simulate the compass gait over any given number of walking cycles.
+
+\section{Active Control Co-Design of the Compass Gait}
+
+In order to perform Active CCD on the compass gait a torque controller was activated at the hip. The addition of this torque controller allowed
+for the problem to be optimized to a more meaningful cost function. Instead of simply optimizing the hip mass of the compass gait, the time per cycle 
+and the total torque used was minimized. The cost function used is shown in \ref{eq:costfunc}, where $\tau_r$ and $t_r$ are reference scalers used 
+to normalize $\tau$ and $t$. The values of $\tau_r$ and $t_r$ which were used are shown in \ref{eq:refvars}
+
+\begin{equation}
+\label{eq:costfunc}
+J = 
+\int_{0}^{t} (\tau / \tau_r)^2 + (t/t_r) \,dt 
+\end{equation}
+
+\begin{equation}
+\label{eq:refvars}
+\tau_r = 6.32 (N*m), 
+t_r = 0.5 (s)
+\end{equation}
+
+Since the hip mass was no longer part of the cost function, it was chosen to also 
+be a design variable of the system. The constraints of the system remained the same, outlined in Figure \ref{fig:OLprob}.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=7cm]{./figures/openloopcontrol.png}
+\caption{The torque and total time to complete a cycle were minimized in the active open loop CCD problem.}
+\label{fig:OLprob}
+\end{figure}
+
+In order to optimize the compass gait through multiple walking cycles and ensure stability, the trajectory generation method used
+in the previous section was used with continued simulation of the system after the first cycle.
+After the trajectory generation was performed and the final states of the system were determined, CCD was performed on the system in order to
+optimize the design variables and the torque over one walking cycle. After the optimal values for the design variables were determined, the system was 
+simulated for a fixed number of cycles with trajectory optimization (torque optimization) according to the set cost function. Figure \ref{fig:OLblock}
+shows a diagram outlining the new trajectory optimization process.
+
+\begin{figure}[h]
+\setlength{\unitlength}{0.14in} % selecting unit length
+\centering % used for centering Figure
+\begin{picture}(32,15) % picture environment with the size (dimensions)
+% 32 length units wide, and 15 units high.
+\put(0,5.5){\vector(1,0){5}}
+\put(5,3.25){\framebox(10,5){Trajectory Generation}}
+\put(20.2,3.25){\framebox(8,5){CCD For 1 Cycle}}
+\put(30.2,3.25){\framebox(12,5){Simulation with Torue Opt.}}
+\put(15,5.5){\vector(1,0){5.2}}
+\put(28.2,5.5){\vector(1,0){2}}
+\put(25.2, 12){\vector(0,-1){3.7}}
+\put(10, 3.25){\vector(0,-1){2}}
+\put(10, 1.25){\vector(1, 0){26.2}}
+\put(36.2, 1.25){\vector(0, 1){2}}
+\put(0,6){Stable IC's}
+\put(15.2, 6){End States}
+\put(15.3, 4.5){$x_1$ and $x_2$}
+\put(23.3, 12.2){New IC's}
+\put(18, 1.5){End States $x_1$ and $x_2$}
+\end{picture}
+\caption{The trajectory generation method used in the previous example was used with continued trajectory optimization after one cycle.} % title of the Figure
+\label{fig:OLblock} % label to refer figure in text
+\end{figure}
+
+In order to compare the compass gaits performance when CCD is used to other compass gait scenarios, three simulations were ran and compared:
+The comass gait with active CCD (Figure \ref{fig:OLblock}), the compass gait with active open loop control without co-design, and the non-optimized passive compass gait system.
+Each of these systems were simulated over 10 cycles and the stability, total torque used, time per cycle, and optimized design variables were compared. Each system used the initial conditions and 
+system parameters shown in Table \ref{tab:params}.
+
+\begin{table}[h]
+\centering
+\caption{Initial Conditions and System Parameters Used to Simulate the Compass Gait.}
+\begin{tabular}{lr}
+\toprule
+Parameter & Value \\
+\midrule
+$m (kg)$ & 5 \\
+$m_H (kg)$ & 10 \\
+$a (m)$ & 0.5 \\
+$b (m)$ & 0.5 \\
+$\gamma (slope, rad)$ & 0.05 \\
+$x_1 (rad)$ & -0.3 \\
+$x_2 (rad)$ & 0.2038 \\
+$x_3 (rad/s)$ & -0.41215 \\
+$x_4 (rad/s)$ & -1.0501 \\
+\end{tabular}
+\label{tab:params}
+\end{table}
+
+Shown in Figures \ref{fig:limcyc_passive} and \ref{fig:timehis_passive} are the limit cycle and total time history of the state variables for the passive compass gait system when simulated using the 
+given initial conditions and system parameters. As can be seen the chosen initial conditions resulted in a stable, well behaved system for the passive compass gait. It can be 
+seen that it takes about three cycles for the system to converge to a single limit cycle; however, there still exists some oscilitations in the limit cycle.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=8cm]{./figures/limitcycle_passivesim.pdf}
+\caption{The limit cycle of the passive compass gait over 10 cycles.}
+\label{fig:limcyc_passive}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=8cm]{./figures/timehis_passivesim.png}
+\caption{The total time history of the state variables of the passive compass gait over 10 cycles.}
+\label{fig:timehis_passive}
+\end{figure}
+
+Similarly, shown in Figures \ref{fig:lim_noco} and \ref{fig:timehis_noco} are the limit cycle and total time history of the state variables and controls for the compass gait 
+using active open loop control without co-design. The same trajectory generation method was used during this simulation, except without co-design (only torque optimization).
+It can be seen that it takes approximataly 6 cycles to converge to a single limit cycle; also, the indivual walking cycles are much more spaced out when compared to the passive system.
+However, the final limit cycle is much smaller than that of the passive limit cycle, resulting in less time per cycle. This can be seen when comparing the total time history of the 
+active system without co-design to the passive system; the average time per cycle and total time of the active system is smaller than that of the passive system.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=8cm]{./figures/limcycle_nocodesign.png}
+\caption{The limit cycle of the active compass gait using open loop control without co-design.}
+\label{fig:lim_noco}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=8cm]{./figures/timehis_nocodesign.png}
+\caption{The total time history of the state variables, control, and cost of the active compass gait without co-design.}
+\label{fig:timehis_noco}
+\end{figure}
+
+Finally, the limit cycle and total time history of the compass gait when CCD is used is shown in Figures \ref{fig:limcyc_CCD} and \ref{fig:timehis_CCD}. When comparing the limit 
+cycle of the CCD compass gait to the other two systems, it can be seen that one small waliking cycle is performed before converging into the final limit cycle in approximataly 4 
+iterations. It can also be seen from the time history that the CCD compass gait has a smaller time per cycle than both the passive system and the open loop system without co-design. 
+It can also be seen that much less torque is used to control the system than the compass gait without co-design.
+
+Shown in Table \ref{tab:opt_vars} are the optimized design variables found using CCD with the compass gait system. As can be seen, in order to reduce torque and time per cycle, 
+the system prefers to have a higher mass at the hip and a longer leg segment below the point mass on the leg. In order to determine the best design variables for a system, 
+the bounds of the design variables should be adjusted as desired.
+
+\begin{table}[h]
+\centering
+\caption{Optimized Design Variables Found Using CCD}
+\begin{tabular}{lr}
+\toprule
+Parameter & Value \\
+\midrule
+$m (kg)$ & 6.45 \\
+$m_H (kg)$ & 100 \\
+$a (m)$ & 1 \\
+$b (m)$ & 0.29 \\
+\end{tabular}
+\label{tab:opt_vars}
+\end{table}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=8cm]{./figures/limcyc_CCD.png}
+\caption{The limit cycle of the compass gait when CCD is used.}
+\label{fig:limcyc_CCD}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=8cm]{./figures/timehis_CCD.png}
+\caption{The total time history of the state variables, control, and cost when CCD is used.}
+\label{fig:timehis_CCD}
+\end{figure}
+
+Shown in Table \ref{tab:time} is the average time per cycle and the total time to complete 10 cycles for each compass gait system which was simulated. As can be seen, the CCD system 
+has the shortest average time per cycle and total time, followed by the open loop system without control, then the passive system. From these results it can be said that 
+CCD is more effective in reducing the time per cycle of the compass gait than control without co-design, and any control at all is much better than a passive system.
+
+\begin{table}[h]
+\centering
+\caption{Average Time per Cycle and Total Time for 10 Cycles of Each Simulation (seconds)}
+\begin{tabular}{ccc}
+\toprule
+Simulation & Average Time/Cycle & Total Time \\
+\midrule
+Passive System & 0.680 & 6.8 \\
+Active System Without Co-Design & 0.594 & 5.94 \\
+System with CCD & 0.531 & 5.31\\
+\end{tabular}
+\label{tab:time}
+\end{table}
+
+The total torque used during the simulations of the compass gait with CCD and the open loop system without co-design can also be compared. Shown in Table \ref{tab:tautot} is 
+the summation of the square of $\tau$ over all 10 cycles, this is shown in equation \ref{eq:summation}. From the table, It can be seen that much more torque is needed to control 
+the non-optimized 
+compass gait when compared to the compass gait with CCD. From the results found in these simulations, it can be concluded that using CCD with the compass gait 
+system can greatly decrease the cost of control when compared to compass gait systems without co-design or optimization.
+
+\begin{equation}
+\label{eq:summation}
+\tau_{sum} = 
+\sum_{t = 0}^{t = t_f}
+\tau^2
+\end{equation}
+
+\begin{table}[h]
+\centering
+\caption{Summation of $\tau^2$ for the CCD and Open Loop Control System}
+\begin{tabular}{lr}
+\toprule
+System & $\sum \tau^2$ \\
+\midrule
+Active System Without Co-Design & 133.46\\
+System with CCD & 29.17\\
+\end{tabular}
+\label{tab:tautot}
+\end{table}
+
+\section{Conclusions and Future Works}
+
+During this study, CCD was applied to the compass gait bipedal robot and the system was optimized first to maximize the mass at the hip with a passive system, 
+then to minimize the torque used and time per cycle with a torque controller at the hip. A trajectory generation method was also used in this paper, 
+where a passive system is first simulated with stable initial conditions and the end states ($x_1$ and $x_2$) are used during the optimization process in order to 
+ensure a stable system.
+
+From the passive optimization, it was shown that not only can optimization 
+of the design variables in the compass gait allow for the hip mass to be maximized, but the optimization can also allow for a greater range of initial conditions 
+which result in a stable first walking cycle. It was shown that when the optimized compass gait was given the same initial conditions which would typically 
+result in an unstable walking cycle with a non-optimized passive system, the system parameters were optimized so that the compass gait had a stable first walking cycle 
+while still maximizing the hip mass.
+
+When CCD was used on the active compass gait, it was shown that the cost of control was reduced by 10.83$\%$ when compared to a system 
+with open loop control without co-design using the same intial conditions and system parameters. Shown in Table \ref{tab:costofcon} is the total cost of control over 10 
+walking cycles when CCD is used compared to active control without co-design. The cost shown is the same cost function shown above in equation \ref{eq:costfunc}. CCD reduced the 
+cost by reducing the torque used and time per cycle in accordance with the cost function. The compass gait with CCD used only 21.8$\%$ as much torque as the system without co-design 
+and the time per cycle was reduced by 0.063 seconds, or 10.6$\%$.
+
+\begin{table}[h]
+\centering
+\caption{Cost of Control for the CCD Compass Gait vs Only Open Loop Control}
+\begin{tabular}{lr}
+\toprule
+System & Cost of Control \\
+\midrule
+Active System Without Co-Design & 11.91\\
+System with CCD & 10.62\\
+Percent Improvement & 10.83$\%$ \\
+\end{tabular}
+\label{tab:costofcon}
+\end{table}
+
+One of the main limitations of this study is that CCD was applied to the compass gait over only one walking cycle using the trajectory generation method,
+then the compass gait was simulated with open-loop 
+control for the given number of cycles.
+Ideally, when using CCD for a system with cyclic behavior it would be desired that the system parameters be optimized based on the entire trajectory of the system 
+(over all cycles); however, due to the structure of the Dymos code, it was only possible to optimize the first walking cycle of the compass gait during this study. 
+In future studies, the compass gait should be optimized over the entire trajectory in order to obtain the best possible system parameters for the system.
+
+Another future study which should be conducted is the use of a closed loop control algorithm in order used to perform CCD. One example of this is using 
+MPC or LQR control in order to optimize and control the compass gait. The use of closed loop CCD with the compass would result in more optimal system parameters 
+and better trajectory optimization.
 

### 2024-05-07 code/compass-gait/compass_torque.py

@@ -127,6 +127,7 @@ class dynamics(om.ExplicitComponent):
         outputs['x3changer'] = x3 - (P11*x3 + P12*x4)
         outputs['x4changer'] = x4 - (P21*x3 + P22*x4)
 
+
     """def compute_partials(self, inputs, partials):
        # computes analytical partials 
 
@@ -203,7 +204,7 @@ class ClosedLoopDynamics(om.Group):
         states_ref = self.options['states_ref']
         # group together dynamics and feedback
         input_names = ['a', 'b', 'x1', 'x2', 'x3', 'x4', 'mh', 'm', 'tau']
-        self.add_subsystem('system_active', dynamics(num_nodes=nn, states_ref=states_ref), promotes_inputs=input_names, promotes_outputs=['*'])
+        self.add_subsystem('system_active', system_active(num_nodes=nn, states_ref=states_ref), promotes_inputs=input_names, promotes_outputs=['*'])
         self.add_subsystem('Feedback', Feedback(num_nodes=nn, states_ref=states_ref), promotes_inputs=['x1', 'x2', 'x3', 'x4', 'K'], promotes_outputs=['tau'])
 
 
@@ -227,9 +228,6 @@ class Feedback(om.ExplicitComponent):
         self.add_output('tau', shape=(nn,), units='N*m')
 
         self.declare_partials(of=['tau'], wrt=['x1', 'x2', 'x3', 'x4'], method='exact', rows=np.arange(nn), cols=np.arange(nn))
-        #self.declare_coloring(wrt=['m_H','m_t','m_s', 'tau',], method='cs', show_summary=False)
-        #self.set_check_partial_options(wrt=['m_H','m_t','m_s', 'tau',], method='fd', step=1e-6)
-
         self.declare_partials(of=['tau'], wrt=['K'], method='cs')
 
     def compute(self, inputs, outputs):
@@ -279,7 +277,7 @@ class CostFunc(om.ExplicitComponent):
         #mh = inputs['mh']
         time = inputs['time']
 
-        outputs['costrate'] = ((tau**2)/40) + (time/0.5)
+        outputs['costrate'] = ((tau**2)/(np.sqrt(40)**2)) + (time/0.5)
         
 
     def compute_partials(self, inputs, partials,):
@@ -288,7 +286,7 @@ class CostFunc(om.ExplicitComponent):
         #mh = inputs['mh']
 
 
-        partials['costrate', 'tau'] = 2*tau/50
[…]

### 2024-05-07 code/compass-gait/lqr_compass.py

@@ -1,0 +1,172 @@
+import numpy as np
+import openmdao.api as om
+import matplotlib.pyplot as plt
+from scipy.linalg import solve_continuous_are
+
+class LQRClosedLoopSystem(om.Group):
+    # given linear system, computes the LQR feedback matrix K, and returns the (components of) closed-loop matrix A_cl where dx/dt = A_cl x
+
+    def initialize(self):
+        self.options.declare('ns', types=int, desc='dimension of state vector x')
+        self.options.declare('nc', types=int, desc='dimension of control vector u')
+        self.options.declare('Q', desc='cost matrix for state, (ns * ns')
+        self.options.declare('R', desc='cost matrix for control, (nc * nc)')
+
+    def setup(self):
+        ns = self.options['ns']
+        nc = self.options['nc']
+
+        # first, solve Riccati equation
+        self.add_subsystem('CARE', Riccati(ns=ns, nc=nc, Q=self.options['Q'], R=self.options['R']), promotes=['*'])
+
+        # feedback matrix K
+        R_inv = np.linalg.solve(self.options['R'], np.eye(nc))
+        k_comp = om.ExecComp('K = dot(dot(Rinv, B.T), P)',
+                             K={'shape' : (nc, ns)},
+                             Rinv={'shape' : (nc, nc), 'val' : R_inv},
+                             B={'shape' : (ns, nc)},
+                             P={'shape' : (ns, ns)})
+        self.add_subsystem('feedback', k_comp, promotes=['*'])
+
+        # closed-loop system matrix. K = R
+        cl_comp = om.ExecComp('A_cl = A - dot(B, K)',
+                              A_cl={'shape' : (ns, ns)},
+                              A={'shape' : (ns, ns)},
+                              B={'shape' : (ns, nc)},
+                              K={'shape' : (nc, ns)})
+        self.add_subsystem('CL', cl_comp, promotes=['*'])
+
+        # convert A_cl matrix to its components. This facilitates the derivative implementation in ODE later
+        ### self.add_subsystem('mat_components', MatrixToComponents2D(), promotes_outputs=['*'])
+        ### self.connect('A_cl', 'mat_components.A')
+
+        # add Newton solver for Riccati eq
[…]

### 2024-05-07 code/compass-gait/run_method2.py

@@ -8,7 +8,7 @@ import numpy as np
 from numpy.linalg import inv
 
 def main():
-    duration_lockphase = 10 # duration of locked knee phase
+    duration_lockphase = 1 # duration of locked knee phase
 
     # defining paramters of the legged walker
 
@@ -26,21 +26,18 @@ def main():
     """
 
     #original initial conditions
-    states_init = [0, 0, 2, -0.4]
+    states_init = {'x1': -0.3, 'x3': -0.41215, 'x2': 0.2038, 'x4': -1.0501} # initial conditions
+    states_final = {'x1': 0.02, 'x3': -0.9, 'x2': -0.1, 'x4': -0.7} # final guess
 
     # number of iterations
-    iterations = 2
-
-    #create plotting arrays
-    x1series = [0]*iterations
-    x2series = [0]*iterations
-    x3series = [0]*iterations
-    x4series = [0]*iterations
+    iterations = 10
 
     for i in range(iterations):
 
-        
-        states_ref = {'x1': states_init[0], 'x3': states_init[1], 'x2': states_init[2], 'x4': states_init[3]}
+        if (i!=0):
+            states_init = {'x1': states_init[0], 'x2': states_init[1], 'x3': states_init[2], 'x4': states_init[3]}
+
+        states_ref = states_init
 
         p = om.Problem()
 
@@ -80,15 +77,15 @@ def main():
         lockphase.add_control('tau', lower = -10, upper = 10, fix_initial=False, units='N*m') # add control torque
 
         # set initial conditions
-        lockphase.add_boundary_constraint('x1', loc='initial', equals=states_init[0])
-        lockphase.add_boundary_constraint('x2', loc='initial', equals=states_init[1])
-        lockphase.add_boundary_constraint('x3', loc='initial', equals=states_init[2])
-        lockphase.add_boundary_constraint('x4', loc='initial', equals=states_init[3])
+        lockphase.add_boundary_constraint('x1', loc='initial', equals=states_init['x1'])
+        lockphase.add_boundary_constraint('x2', loc='initial', equals=states_init['x2'])
+        lockphase.add_boundary_constraint('x3', loc='initial', equals=states_init['x3'])
+        lockphase.add_boundary_constraint('x4', loc='initial', equals=states_init['x4'])
 
         # add paramaters
[…]

### 2024-05-07 code/compass-gait/runcompass.py

@@ -14,8 +14,6 @@ def main(co_design=True):
 
     a = 0.5
     b = 0.5
-    body_mass = 6
-    carrier_mass = 3
     mh = 10
     m = 5
     l = a + b
@@ -24,10 +22,8 @@ def main(co_design=True):
 
     density = 5
 
-    """ 
-    walker will complete one full cycle -- states will be the same at the end as they were at the beginning
-    """
-    states_init = {'x1': -0.3, 'x3': -0.41215, 'x2': 0.2038, 'x4': -1.0501} # initial conditions
+
+    states_init = {'x1': -0.23, 'x3': -0.4, 'x2': 0.25, 'x4': -0.75}
     states_final = {'x1': 0.0264, 'x3': -0.895, 'x2': -0.1264, 'x4': -0.669} # final guess
     #states_final = {'x1': .1, 'x3': 0, 'x2': -0.2, 'x4': 0}
 
@@ -94,20 +90,19 @@ def main(co_design=True):
     # paramaters - same for both phases
     initphase.add_parameter('a', val=a, units='m', static_target=True)
     initphase.add_parameter('b', val=b, units='m', static_target=True)
-    initphase.add_parameter('mh', opt=True, val=mh, lower = 5, upper=80, units='kg', static_target=True)
+    initphase.add_parameter('mh', opt=co_design, val=mh, lower = 5, upper=1000, units='kg', static_target=True)
     initphase.add_parameter('m', val=m, units='kg', static_target=True)
 
     # transition boudary contraints
     initphase.add_boundary_constraint('phi_bounds', loc='final', equals=phi_contraint,  units='rad')
 
-    # set design var a to be a and b
-    p.model.connect('a', 'traj.initphase.parameters:a')
-    p.model.connect('b', 'traj.initphase.parameters:b')
-    p.model.connect('m', 'traj.initphase.parameters:m')
+    if co_design:
+        # set design var a to be a and b
+        p.model.connect('a', 'traj.initphase.parameters:a')
+        p.model.connect('b', 'traj.initphase.parameters:b')
+        p.model.connect('m', 'traj.initphase.parameters:m')
 
-    initphase.add_timeseries_output('costrate', output_name='costrate')
-
-    initphase.add_objective('costrate')
+    initphase.add_objective('mh', scaler=-1)
 
     p.setup(check=True)
 
@@ -126,9 +121,7 @@ def main(co_design=True):
 
     #om.n2(p)
 
-    # print cost
[…]

### 2024-05-07 code/compass-gait/runwithsim_OL.py

@@ -234,9 +234,9 @@ def traj_opt(a=0.5, b=0.5, mh=10, m=5, states_final=[0, 0, 0, 0], states_init=[0
     phase.set_time_options(fix_initial=True, initial_val=0, fix_duration=False, units='s') # set time options for simulation
 
     #states for init phase
-    phase.add_state('x1', fix_initial=True, fix_final=True, rate_source='x1_dot', units='rad')
-    phase.add_state('x3', fix_initial=True, rate_source='x3_dot', units='rad/s')
-    phase.add_state('x2',  fix_initial=True, fix_final=True, rate_source='x2_dot', units='rad')
+    phase.add_state('x1', fix_initial=True, upper=1, lower=-1, fix_final=True, rate_source='x1_dot', units='rad')
+    phase.add_state('x3', fix_initial=True,  rate_source='x3_dot', units='rad/s')
+    phase.add_state('x2',  fix_initial=True, upper=1, lower=-1, fix_final=True, rate_source='x2_dot', units='rad')
     phase.add_state('x4', fix_initial=True, rate_source='x4_dot', units='rad/s')
     phase.add_state('cost', fix_initial=True, rate_source='costrate')
 
@@ -313,7 +313,7 @@ def traj_opt(a=0.5, b=0.5, mh=10, m=5, states_final=[0, 0, 0, 0], states_init=[0
 if __name__ == '__main__':
     states_final = traj_gen() # 1) run trajectory generation and get final states
 
-    first_states_init = [-0.3, 0.2038, -0.41215, -1.05] # in order : x1, x2, x3, x4 - default IC's
+    first_states_init =[-0.3, 0.2038, -0.41215, -1.0501] # in order : x1, x2, x3, x4 - default IC's
     #first_states_init = [0, 0, 0, 0]
     p2, _ = traj_opt(states_final=states_final, states_init=first_states_init, co_design=True) # 2) run optimization with co-designand torque to get sys paramaters and first cycle
     a = p2.get_val('a', units='m')
@@ -328,13 +328,15 @@ if __name__ == '__main__':
     timearr = p2.get_val('traj.phase.timeseries.time')
     tauarr = p2.get_val('traj.phase.timeseries.controls:tau')
     costarr = p2.get_val('traj.phase.timeseries.states:cost')
+    costpercyc = costarr[-1]
 
     num_sin_points = len(x1arr)
+    endtimes = timearr[-1]
[…]

### 2024-05-07 code/compass-gait/runwithsim_passive.py

@@ -2,14 +2,13 @@ import numpy as np
 import openmdao.api as om
 import dymos as dm
 from compass_notorque import system_passive
-from compass_torque import ClosedLoopDynamics
 import matplotlib.pyplot as plt
 from dymos.examples.plotting import plot_results
 import numpy as np
 from numpy.linalg import inv
 
 """
-CLOSED LOOP CONTROL OVER LIMIT CYCLE (CO-DESIGN FOR FIRST CYCLE ONLY) WITH OBJ: MINIMIZE TIME AND TAU WITH TRAJECTORY GENERATION
+OPEN LOOP CO DESIGN OVER 1 CYCLE - PASSIVE OPTIMIZATION PROBLEM
 """
 
 # defining paramters of the legged walker
@@ -20,9 +19,6 @@ m = 5
 l = a + b
 phi = 0.05 # slope - 3 degrees
 phi_contraint = -2*phi
-
-
-
 
 def traj_gen():
     duration_lockphase = 1 # duration of locked knee phase
@@ -217,15 +213,12 @@ def traj_opt(a=0.5, b=0.5, mh=10, m=5, states_final=[0, 0, 0, 0], states_init=[0
     design_vars.add_output('b', val=b, units='m')
     design_vars.add_output('mh', val=mh, units='kg')
     design_vars.add_output('density', val=density, units='kg/m')
-    #design_vars.add_output('K', val=np.zeros((1, 4)))
-
-    #p2.model.add_design_var('K', ref=np.zeros((1, 4)))
 
     if co_design:
         # add design var a 
         p2.model.add_design_var('a', units='m', lower=0.1, upper=1,  ref=0.5)
         p2.model.add_design_var('b', units='m', lower=0.1, upper=1,  ref=0.5)
-        p2.model.add_design_var('mh', units='kg', lower=5, upper=100)
+        p2.model.add_design_var('mh', units='kg', lower=5, upper=80)
 
         leg_mass_comp = om.ExecComp('m=density*(a+b)')
         leg_mass = p2.model.add_subsystem('m', leg_mass_comp, promotes=['*'])
@@ -233,7 +226,7 @@ def traj_opt(a=0.5, b=0.5, mh=10, m=5, states_final=[0, 0, 0, 0], states_init=[0
     traj = p2.model.add_subsystem('traj', dm.Trajectory()) # init trajectory
 
     #add init phase
[…]

### 2024-05-07 R0_journal/main.tex
paper update

@@
-\documentclass[smallextended]{./springer/svjour3}
+\documentclass{./springer/svjour3}
 \usepackage{graphicx} % This lets you include figures
@@
 as calculate the updated angular velocities after heel strike are shown in \ref{eq:trans1} and \ref{eq:trans2}, where $x^+$ represents the state
-variable after heel strike and $x^-$ represents the state variable before impact, and $\mb{H}$ is the angular velocity transition matrix shown in the appendix.
+variable after heel strike and $x^-$ represents the state variable before impact, $\mb{Q^-}$ and $\mb{Q^+}$ are the angular velocity transition matrices shown in the appendix.
 
@@
 \label{eq:trans2}
+\mb{Q^+}
 \begin{bmatrix}
@@
  = 
-\mb{H}
+\mb{Q^-}
 \begin{bmatrix}
@@
 \centering
-\includegraphics[width=9cm]{./figures/onecycle.pdf}
+\includegraphics[width=10cm]{./figures/onecycle.pdf}
 \caption{One walking limit cycle of the compass gait. The swing leg makes a large swing (shown in blue, a large cicle) around while the stance leg only pivots (shown in orange.)}
@@
 \centering
-\includegraphics[width=10cm]{./figures/states_onecycle.pdf}
+\includegraphics[width=9cm]{./figures/states_onecycle.pdf}
 \caption{States of the compas gait over time for one walking cycle.}
@@
 \end{figure}
+
+\clearpage
 
@@
 and simulate the compass gait over any given number of walking cycles.
+
+\clearpage
 
@@
 
+\clearpage
+\appendix
+
+\section{Compass Gait Equations}
+
+Shown in this appendix is a more complete version of the dynamics equations of the compass gait. The swing stage dynamics can be written in standard planar manipulator 
+dynamics form, shown in \ref{eq:swingeq}. The $\mb{M}$, $\mb{N}$, $\mb{G}$, and $\mb{U}$ matrices are also shown in \ref{eq:M}, \ref{eq:N}, \ref{eq:G}, and \ref{eq:U} respectively.
+
+\begin{equation}
+\label{eq:swingeq}
+\mb{M(X)}\mb{\ddot{X}} + \mb{N(X, \dot{X})}\mb{\dot{X}} + \mb{G(X)} = \tau\mb{U}
+\end{equation}
+
+\begin{equation}
+\mb{M} = 
+\label{eq:M}
+\begin{bmatrix}
+\\mb^2 &  -mlbcos(x_2 - x_1)\\
+\\-mlbcos(x_2 - x_1) &  m_{H}l^2 + m(l^2 + a^2)\\
+\end{bmatrix}
+\end{equation}
+
+\begin{equation}
+\mb{N} = 
+\label{eq:N}
+\begin{bmatrix}
+\\0 &  mlb\dot{x_2}sin(x_2 - x_1)\\
+\\-mlb\dot{x_1}sin(x_2 - x_1) &  0\\
+\end{bmatrix}
+\end{equation}
+
+\begin{equation}
+\mb{G} = 
+\label{eq:G}
+\begin{bmatrix}
+\\mgbsin(x_1)\\
+\\-(m_{H}l + m(a+l))gsin(x_2)\\
+\end{bmatrix}
+\end{equation}
+
+\begin{equation}
+\mb{U} = 
+\label{eq:U}
+\begin{bmatrix}
+\\1\\
+\\-1\\
+\\0\\
+\\0\\
+\end{bmatrix}
+\end{equation}
+
+The transition equations when heel strike occurs are shown in \ref{eq:transeqn} and \ref{eq:transeqn2}, where $x^+$ signifies a state variables after heel strike and $x^-$ 
+signals a state variable before
+heel strike. The transition matrices $\mb{Q^+}$ and $\mb{Q^-}$ are shown in \ref{eq:Q-} and \ref{}. Note that $\alpha$ represents half the angle between the swing and stance 
+legs.
+
+\begin{equation}
+\label{eq:transeqn}
+\begin{bmatrix}
+\\x_1^+\\
+\\x_2^+\\
+\end{bmatrix}
+ = 
+\begin{bmatrix}
+0 & 1\\
+1 & 0\\
+\end{bmatrix}
+\begin{bmatrix}
+\\x_1^-\\
+\\x_2^-\\
+\end{bmatrix}
+\end{equation}
+
+\begin{equation}
+\label{eq:transeqn2}
+\mb{Q^+}
+\begin{bmatrix}
+\\x_3^+\\
+\\x_4^+\\
+\end{bmatrix}
+= 
+\mb{Q^-}
+\begin{bmatrix}
+\\x_3^-\\
+\\x_4^-\\
+\end{bmatrix}
+\end{equation}
+
+\begin{equation}
+\mb{Q^-} = 
+\label{eq:Q-}
+\begin{bmatrix}
+\\-mab & -mab + (m_Hl^2+2mal)cos(2\alpha)\\
+\\0 & -mab\\
+\end{bmatrix}
+\end{equation}
+
+\begin{equation}
+\mb{Q^+} = 
+\label{eq:Q+}
+\begin{bmatrix}
+\\mb(b-lcos(2\alpha)) & ml(l-bcos(2\alpha)) + ma^2 + m_Hl^2\\
+\\mb^2 & -mblcos(2\alpha)\\
+\end{bmatrix}
+\end{equation}
+
 \end{document}

### 2024-05-15 R0_journal/main.tex
bib

@@
 
-\title{Research project title}
+\title{Optimization of the Walking Compass Gait Robot With the Use of Control Co-Design}
 \author{Max Howell}
@@
 \begin{document}
-\bibliographystyle{plain}
-\bibliography{references}
 
@@
 when a system is designed an iterative process is used to determine the best parameters for the system; however, this can be time consuming and inefficient. 
-It has been shown in \cite{} that the cost of control can be decreased by atleast xx\% in systems such as a cartpole pendulum and a quadrotor by using CCD. CCD applications range
-from design optimizatoin of (examples of CCD being used with citations)
+It has been shown in \cite{wu2022control} that CCD can be used to improve the weight reduction of a high-acceleration precision motion system by 42$\%$, and 
+the bandwith of the system by 28$\%$. It has also been shown in \cite{Mandali2023} that CCD can be used to control other non-mechancial such as thermal 
+management systems.
 
@@
 \centering
-\includegraphics[width=10cm]{./figures/compassgaitmodel.png}
+\includegraphics[width=8cm]{./figures/compassgaitmodel.png}
 \caption{Schematic of the compass gait model. Angles $\theta_{sw}$ and $\theta_{st}$ are shown as the swing leg angle and stance leg angle, respectively.}
@@
 
-The dynamic equations of the swing stage of the compass gait, from \cite{french confrence}, can be written in standard planar manupular dynamics form. 
+The dynamic equations of the swing stage of the compass gait, from \cite{goswami:inria-00073701}, can be written in standard planar manupular dynamics form. 
 This is shown in \ref{eq:swingeqn}, where $\mb{M}$, $\mb{N}$, and $\mb{G}$ are matrices shown in the appendix, and $\mb{U}$
@@
 In order to acheive co-design with a simple passive compass gait system, an optimization problem was proposed.
-This optimization problem, outlined in Figure \ref{fig:optsetup}, involved maximizing the point mass at the hip ($m_H$) in order to simulate the
+This optimization problem, outlined in  \ref{eq:optprobsetup}, involved maximizing the point mass at the hip ($m_H$) in order to simulate the
 compass gait robot "carrying" a heavy object. In order to acheive this goal, the lengths below and above the leg point mass, a and b, were optimized.
@@
 
-\begin{figure}[!h]
-\centering
-\includegraphics[width=6cm]{./figures/opt1prob.png}
-\caption{Outline of the simple passive optimization problem. The goal is to incrase the range of stable initial conditions while maximizing the applied hip mass.}
-\label{fig:optsetup}
-\end{figure}
+\begin{equation}
+\begin{aligned}
+\text{Objective: Maximize $m_H$}\\
+\text{Subject to:}\\
+\text{Dynamic Constraints of System}\\
+\text{Design Variables: 0.1$<$a$<$1, 0.1$<$b$<$1, (m)}\\
+\text{Hip Mass: 5$<$$m_H$$<$80, (kg)}\\
+\text{Leg Mass: } m = 5(a+b), \text{(kg)}\\
+\text{End States $x_1$ and $x_2$ are Constrained Based on Trajectory Generation}
+\end{aligned}
+\label{eq:optprobsetup}
+\end{equation}
 
@@
 \centering
-\caption{Initial Conditions and System Parameters of the Non-Optimized Compass Gait Shown in \ref{fig:nonoptlimcycle}}
+\caption{Optimized System Parameters of the compass gait.}
 \begin{tabular}{lr}
@@
 Since the hip mass was no longer part of the cost function, it was chosen to also 
-be a design variable of the system. The constraints of the system remained the same, outlined in Figure \ref{fig:OLprob}.
+be a design variable of the system. The constraints of the system remained the same, outlined in \ref{eq:OLprob}.
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=7cm]{./figures/openloopcontrol.png}
-\caption{The torque and total time to complete a cycle were minimized in the active open loop CCD problem.}
-\label{fig:OLprob}
-\end{figure}
+\begin{equation}
+\begin{aligned}
+\text{Minimize: } J = \int_{0}^{t} (\tau / \tau_r)^2 + (t/t_r) \,dt \\
+\text{Subject to:}\\
+\text{Dynamic Constraints of System}\\
+\text{Design Variables: } 0.1<a<1 (m), 0.1<b<1 (m), 0<m_H<100 \text{kg}\\
+\text{Leg Mass: } m = 5(a+b)\\
+\text{End States of $x_1$ and $x_2$ Constrained by Trajectory Generation}
+\end{aligned}
+\label{eq:OLprob}
+\end{equation}
 
@@
 \caption{Average Time per Cycle and Total Time for 10 Cycles of Each Simulation (seconds)}
-\begin{tabular}{ccc}
+\begin{tabular}{cccc}
 \toprule
-Simulation & Average Time/Cycle & Total Time \\
+Simulation & Average Time/Cycle & Total Time & Percent Improvement\\
 \midrule
-Passive System & 0.680 & 6.8 \\
-Active System Without Co-Design & 0.594 & 5.94 \\
-System with CCD & 0.531 & 5.31\\
+Passive System & 0.680 & 6.8 & N/A\\
+Active System Without Co-Design & 0.594 & 5.94 & 12.66$\%$ \\
+System with CCD & 0.531 & 5.31 & 21.91$\%$\\
 \end{tabular}
@@
 \caption{Summation of $\tau^2$ for the CCD and Open Loop Control System}
-\begin{tabular}{lr}
+\begin{tabular}{ccc}
 \toprule
-System & $\sum \tau^2$ \\
+System & $\sum \tau^2$ & Percent Improvement \\
 \midrule
-Active System Without Co-Design & 133.46\\
-System with CCD & 29.17\\
+Active System Without Co-Design & 133.46 & N/A\\
+System with CCD & 29.17 & 78.1$\%$\\
 \end{tabular}
@@
 Shown in this appendix is a more complete version of the dynamics equations of the compass gait. The swing stage dynamics can be written in standard planar manipulator 
-dynamics form, shown in \ref{eq:swingeq}. The $\mb{M}$, $\mb{N}$, $\mb{G}$, and $\mb{U}$ matrices are also shown in \ref{eq:M}, \ref{eq:N}, \ref{eq:G}, and \ref{eq:U} respectively.
+dynamics form, shown in \ref{eq:swingeq}. The $\mb{M}$, $\mb{N}$, $\mb{G}$, and $\mb{U2}$ matrices are also shown in \ref{eq:M}, \ref{eq:N}, \ref{eq:G}, and \ref{eq:U} respectively.
 
@@
 \mb{U} = 
-\label{eq:U}
+\label{eq:U2}
 \begin{bmatrix}
@@
 
+\bibliography{bib/references}
+\bibliographystyle{plain}
+
 \end{document}

### 2024-05-15 R0_journal/main.tex
...

@@
 \begin{document}
-
 \maketitle{}
-
+\bibliographystyle{ieeetr}
 \begin{abstract}
@@
 \centering
-\includegraphics[width=10cm]{./figures/onecycle.pdf}
+\includegraphics[width=6.5cm]{./figures/onecycle.pdf}
 \caption{One walking limit cycle of the compass gait. The swing leg makes a large swing (shown in blue, a large cicle) around while the stance leg only pivots (shown in orange.)}
@@
 \centering
-\includegraphics[width=9cm]{./figures/states_onecycle.pdf}
+\includegraphics[width=8cm]{./figures/states_onecycle.pdf}
 \caption{States of the compas gait over time for one walking cycle.}
@@
 \centering
-\includegraphics[width=10cm]{./figures/unstable_onecycle.pdf}
+\includegraphics[width=8cm]{./figures/unstable_onecycle.pdf}
 \caption{An example of an unstable limit cycle. In this case, the initial angular velocities make it impossible for the compass gait to start a walking cycle.}
@@
 \begin{aligned}
-\text{Objective: Maximize $m_H$}\\
-\text{Subject to:}\\
-\text{Dynamic Constraints of System}\\
-\text{Design Variables: 0.1$<$a$<$1, 0.1$<$b$<$1, (m)}\\
-\text{Hip Mass: 5$<$$m_H$$<$80, (kg)}\\
-\text{Leg Mass: } m = 5(a+b), \text{(kg)}\\
-\text{End States $x_1$ and $x_2$ are Constrained Based on Trajectory Generation}
+&\text{Objective: Maximize $m_H$}\\
+&\text{Subject to:}\\
+&\text{Dynamic Constraints of System}\\
+&\text{Design Variables: 0.1$<$a$<$1, 0.1$<$b$<$1, (m)}\\
+&\text{Hip Mass: 5$<$$m_H$$<$80, (kg)}\\
+&\text{Leg Mass: } m = 5(a+b), \text{(kg)}\\
+&\text{End States $x_1$ and $x_2$ are Constrained by Trajectory Generation}
 \end{aligned}
@@
 \begin{aligned}
-\text{Minimize: } J = \int_{0}^{t} (\tau / \tau_r)^2 + (t/t_r) \,dt \\
-\text{Subject to:}\\
-\text{Dynamic Constraints of System}\\
-\text{Design Variables: } 0.1<a<1 (m), 0.1<b<1 (m), 0<m_H<100 \text{kg}\\
-\text{Leg Mass: } m = 5(a+b)\\
-\text{End States of $x_1$ and $x_2$ Constrained by Trajectory Generation}
+&\text{Minimize: } J = \int_{0}^{t} (\tau / \tau_r)^2 + (t/t_r) \,dt \\
+&\text{Subject to:}\\
+&\text{Dynamic Constraints of System}\\
+&\text{Design Variables: } 0.1<a<1 (m), 0.1<b<1 (m), 0<m_H<100 \text{kg}\\
+&\text{Leg Mass: } m = 5(a+b)\\
+&\text{End States of $x_1$ and $x_2$ Constrained by Trajectory Generation}
 \end{aligned}
@@
 Shown in this appendix is a more complete version of the dynamics equations of the compass gait. The swing stage dynamics can be written in standard planar manipulator 
-dynamics form, shown in \ref{eq:swingeq}. The $\mb{M}$, $\mb{N}$, $\mb{G}$, and $\mb{U2}$ matrices are also shown in \ref{eq:M}, \ref{eq:N}, \ref{eq:G}, and \ref{eq:U} respectively.
+dynamics form, shown in \ref{eq:swingeq}. The $\mb{M}$, $\mb{N}$, $\mb{G}$, and $\mb{U}$ matrices are also shown in \ref{eq:M}, \ref{eq:N}, \ref{eq:G}, and \ref{eq:U2} respectively.
 
@@
 signals a state variable before
-heel strike. The transition matrices $\mb{Q^+}$ and $\mb{Q^-}$ are shown in \ref{eq:Q-} and \ref{}. Note that $\alpha$ represents half the angle between the swing and stance 
+heel strike. The transition matrices $\mb{Q^-}$ and $\mb{Q^+}$ are shown in \ref{eq:Q-} and \ref{eq:Q+}. Note that $\alpha$ represents half the angle between the swing and stance 
 legs.
@@
 
+\section*{Acknowledgement}
+\LARGE
+I would like to thank my professor, Sicheng He, who made this project possible. The knowledge and support he provided 
+helped guide me through this project. I would also like to thank Shugo Kaneko for the help he provided by debugging my Dymos code which 
+signficantly sped up the process of developing the simulation of the compass gait.
+
+\clearpage
+
+\pagestyle{plain}
+\normalsize
 \bibliography{bib/references}
-\bibliographystyle{plain}
 

### 2024-05-15 code/compass-gait/run_method2.py

@@ -30,7 +30,7 @@ def main():
     states_final = {'x1': 0.02, 'x3': -0.9, 'x2': -0.1, 'x4': -0.7} # final guess
 
     # number of iterations
-    iterations = 10
+    iterations = 3
 
     for i in range(iterations):
 
@@ -194,7 +194,7 @@ def main():
     # animate motion
     from animate import animate_compass
     num_points = len(x1arr)
-    animate_compass(x1arr.reshape(num_points), x2arr.reshape(num_points), a, b, phi, saveFig=True, name='runmultpassive.gif',gif_fps=40, iter=iterations, num_iter_points=num_iter)
+    animate_compass(x1arr.reshape(num_points), x2arr.reshape(num_points), a, b, phi, saveFig=True, name='runmultpassive.gif',gif_fps=29, iter=iterations, num_iter_points=num_iter)
 
     fig, (ax1, ax2, ax3, ax4) = plt.subplots(4)
     fig.suptitle('States Over Entire Range - Passive System')

### 2024-05-15 code/compass-gait/runwithsim_OL.py

@@ -336,7 +336,7 @@ if __name__ == '__main__':
 
 
     # 3) loop through traj optimization WITHOUT co design using found parameters to simulate system with open loop torque control
-    iterations = 9 # number of loops
+    iterations = 2 # number of loops
 
     for i in range(iterations):
         
@@ -407,7 +407,7 @@ if __name__ == '__main__':
     num_points = len(x1arr)
 
     from animate import animate_compass
-    animate_compass(x1arr.reshape(num_points), x2arr.reshape(num_points), a, b, phi, name='openloop_cocontrol.gif', saveFig=True, gif_fps=40, iter=iterations+1, num_iter_points=num_sin_points)
+    animate_compass(x1arr.reshape(num_points), x2arr.reshape(num_points), a, b, phi, name='openloop_cocontrol.gif', saveFig=True, gif_fps=42, iter=iterations+1, num_iter_points=num_sin_points)
 
     # plot results
     fig, (ax1, ax2, ax3, ax4, ax5, ax6) = plt.subplots(6)

### 2024-12-16 code/nl_mpc_example/dompc/cartpole_mujoco/codesign.py
minor

@@ -1,0 +1,3 @@
+import numpy as np
+from main import balance
+

### 2024-12-16 code/nl_mpc_example/dompc/cartpole_mujoco/main.py

@@ -10,142 +10,147 @@ sys.path.append(rel_do_mpc_path)
 
 # Import do_mpc package:
 import do_mpc
-delta_t = .02
-
 
 # import model and controller
 from model import model_set
 from controller import control
 
-# initialize mujoco and renderer
-from mj_interface import mjmod_init, mjrend_init
-x0 = [0, np.deg2rad(180)]
-model, data = mjmod_init(x0)
-window, camera, scene, context, viewport = mjrend_init(model, data)
-   
+# import mujoco interface
+from mj_interface import mjmod_init, mjrend_init, linearize
 
-from mj_interface import linearize
-
-# set matrices for plotting
-xarr = []
-thetaarr = []
-farr = []
-#jarr = []
-tarr = []
-yarr = []
-the_dmpc = []
-
-
-# start main loop
-x = np.zeros(4)
-step = 1
-rgb = []
-depth = []
-
-# set up recording
+# import opengl to save results as a mp4
 import cv2
 from OpenGL.GL import *
 from OpenGL.GLU import *
 from OpenGL.GLUT import *
-width, height = glfw.get_framebuffer_size(window)
-fourcc = cv2.VideoWriter_fourcc(*'mp4v')
-video_writer = cv2.VideoWriter('cartpole.mp4', fourcc, 30.0, (width, height))
 
-while(not glfw.window_should_close(window)):
+# make function to do co design using finite difference
+def balance(delta_t = 0.02, plotting = False):
 
-    # record image from window and save to video:
-    glReadBuffer(GL_FRONT)
-    pixel_data = glReadPixels(0, 0, width, height, GL_BGR, GL_UNSIGNED_BYTE)
-    # Convert pixel data to a numpy array
-    image = np.frombuffer(pixel_data, dtype=np.uint8)
-    image = image.reshape(height, width, 3)
-    # Flip the image vertically (OpenGL's origin is at the bottom left)
-    image = np.flipud(image)
-    video_writer.write(image)
+    
+    x0 = [0, np.deg2rad(180)]
+    model, data = mjmod_init(x0)
 
-    # mj step 1: pre control
-    mujoco.mj_step1(model, data)
+    window, camera, scene, context, viewport = mjrend_init(model, data)
+    
 
-    # get linearized system
-    A, B = linearize(model, data)
-    #print(A)
-    #print(B)
-    # model and controller
-    dmpc_mod = model_set(A, B)
[…]

### 2024-12-17 code/nl_mpc_example/dompc/cartpole_mujoco/codesign.py
working still

@@ -1,3 +1,44 @@
 import numpy as np
 from main import balance
 
+# do control co-design to optimize the length of the 
+# pole in cart pole system to minimize force used to balance
+
+cost = 25000
+cost_prev = 0
+step = 0.1
+thres = 5000
+psize_cur = 0.3
+psize_prev = 0
+
+iter = 1
+while (cost > thres):
+
+    if iter != 1: # ignore if first iteration
+        diff_cost = cost - cost_prev
+        diff_size = psize_cur - psize_prev
+        psize_prev = psize_cur
+        if diff_cost > 0: # if cost went up with last step
+            if diff_size > 0: # and if pole size increased
+                psize_cur =- step # decrease pole size for next step
+            else: # and if pole size decreased
+                psize_cur =+ step #increase pole size for next step
+        else: # if cost went down
+            if diff_size > 0: # and if pole size increased
+                psize_cur =+ step # decrease pole size for next step
+            else: # and if pole size decreased
+                psize_cur =- step #increase pole size for next step
+    
+    cost_prev = cost
+
+    cost, tf = balance(polelen = psize_cur)
+    print(cost, psize_cur, iter)
+    iter += 1
+    
+
+print(f'Reached Threshhold Cost of {cost} with pole size of {psize_cur}')
+
+        
+
+
+    

### 2024-12-17 code/nl_mpc_example/dompc/cartpole_mujoco/main.py

@@ -16,7 +16,7 @@ from model import model_set
 from controller import control
 
 # import mujoco interface
-from mj_interface import mjmod_init, mjrend_init, linearize
+from mj_interface import mjmod_init, mjrend_init, linearize, setpolelen
 
 # import opengl to save results as a mp4
 import cv2
@@ -25,15 +25,17 @@ from OpenGL.GLU import *
 from OpenGL.GLUT import *
 
 # make function to do co design using finite difference
-def balance(delta_t = 0.02, plotting = False):
+def balance(delta_t = 0.02, plotting = False, polelen = .3):
 
-    
+    # set initial conditions
     x0 = [0, np.deg2rad(180)]
     model, data = mjmod_init(x0)
 
-    window, camera, scene, context, viewport = mjrend_init(model, data)
-    
+    # set pole length
+    setpolelen(model, data, polelen)
 
+    # init window
+    window, camera, scene, context, viewport = mjrend_init(model, data)
     
     # set matrices for plotting
     xarr = []
@@ -44,6 +46,7 @@ def balance(delta_t = 0.02, plotting = False):
     yarr = []
     the_dmpc = []
 
+    cost = 0 # initialize cost 
 
     # start main loop
     x = np.zeros(4)
@@ -51,13 +54,26 @@ def balance(delta_t = 0.02, plotting = False):
     rgb = []
     depth = []
 
+    xtol = .1
+    thetatol = .05 # 3 degrees
+    dxtol = 0.1
+    dthetatol = .05
+
     # set up recording
     if plotting:
         width, height = glfw.get_framebuffer_size(window)
         fourcc = cv2.VideoWriter_fourcc(*'mp4v')
         video_writer = cv2.VideoWriter('cartpole.mp4', fourcc, 30.0, (width, height))
 
-    while(not glfw.window_should_close(window)):
+    stop = False
+    while(stop==False):
+        if step != 1:
+            if abs(curx) < xtol and abs(curtheta) < thetatol and abs(curdx) < dxtol and abs(curdtheta) < dthetatol:
+                stop = True
+            elif glfw.window_should_close(window):
+                stop = True
+
+
 
         if plotting:
             # record image from window and save to video:
@@ -107,6 +123,8 @@ def balance(delta_t = 0.02, plotting = False):
         cury = y_next[0]
         curthe_dmpc = y_next[1]
 
+        cost += u**2
+
 
[…]

### 2024-12-17 code/nl_mpc_example/dompc/cartpole_mujoco/mj_interface.py

@@ -44,7 +44,7 @@ def mjrend_init(model, data):
     scene = mujoco.MjvScene(model, 6000)
     camera = mujoco.MjvCamera()
     camera.trackbodyid = 2
-    camera.distance = 3
+    camera.distance = 8
     camera.azimuth = 90
     camera.elevation = -20
     mujoco.mjv_updateScene(
@@ -131,5 +131,10 @@ def linearize(model, data):
             B[k,i] = (xdot[k]-xdot0[k])/pert
 
     return A,B
+
+def setpolelen(model, data, leng=0.6):
+    model.geom_size[2,1] = leng
+    model.geom_pos[2,2] = leng
+    model.body_ipos[2,2] = leng
 
 

### 2024-12-17 code/nl_mpc_example/dompc/cartpole_mujoco/controller.py
slightly better

@@ -34,7 +34,7 @@ def control (model, delta_t):
 
     mpc.set_objective(mterm=mterm, lterm=lterm)
 
-    mpc.set_rterm(u=1e-2) # input penalty
+    mpc.set_rterm(u=2e-2) # input penalty
 
     max_x = np.array([[1], [100], [100], [100]])
     min_x = np.array([[-1], [-100], [-100], [-100]])

### 2024-12-17 code/nl_mpc_example/dompc/cartpole_mujoco/main.py

@@ -179,4 +179,4 @@ def balance(delta_t = 0.02, plotting = False, polelen = .3):
 
 
 if __name__ == "__main__":
-    balance(polelen=0.3, plotting=True)
+    balance(polelen=0.25, plotting=True)

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_mujoco/codesign.py
...

@@ -5,38 +5,47 @@ from main import balance
 # pole in cart pole system to minimize force used to balance
 
 cost = 25000
-cost_prev = 0
+cost_prev = 999999
 step = 0.1
-thres = 5000
-psize_cur = 0.3
+thres = 4500
+psize_cur = 4
 psize_prev = 0
 
 iter = 1
+stop = False
 while (cost > thres):
-
+    if stop:
+        break
     if iter != 1: # ignore if first iteration
         diff_cost = cost - cost_prev
         diff_size = psize_cur - psize_prev
         psize_prev = psize_cur
+        if Fail:
+            stop = True
+            if diff_size > 0:
+                psize_cur -= step
+            else:
+                psize_cur += step
+
         if diff_cost > 0: # if cost went up with last step
             if diff_size > 0: # and if pole size increased
-                psize_cur =- step # decrease pole size for next step
+                psize_cur -= step # decrease pole size for next step
             else: # and if pole size decreased
-                psize_cur =+ step #increase pole size for next step
+                psize_cur += step #increase pole size for next step
         else: # if cost went down
             if diff_size > 0: # and if pole size increased
-                psize_cur =+ step # decrease pole size for next step
+                psize_cur += step # decrease pole size for next step
             else: # and if pole size decreased
-                psize_cur =- step #increase pole size for next step
+                psize_cur -= step #increase pole size for next step
     
     cost_prev = cost
 
-    cost, tf = balance(polelen = psize_cur)
+    cost, tf, Fail = balance(cart_m = psize_cur)
     print(cost, psize_cur, iter)
     iter += 1
     
 
-print(f'Reached Threshhold Cost of {cost} with pole size of {psize_cur}')
+print(f'Reached Threshhold Cost of {cost} with parameter of {psize_cur}')
 
         
 

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_mujoco/controller.py

@@ -34,7 +34,7 @@ def control (model, delta_t):
 
     mpc.set_objective(mterm=mterm, lterm=lterm)
 
-    mpc.set_rterm(u=2e-2) # input penalty
+    mpc.set_rterm(u=0) # input penalty
 
     max_x = np.array([[1], [100], [100], [100]])
     min_x = np.array([[-1], [-100], [-100], [-100]])
@@ -46,10 +46,10 @@ def control (model, delta_t):
     mpc.bounds['upper','_x','x'] = max_x
 
     # lower bounds of the input
-    mpc.bounds['lower','_u','u'] = -20
+    mpc.bounds['lower','_u','u'] = -30
 
     # upper bounds of the input
-    mpc.bounds['upper','_u','u'] =  20
+    mpc.bounds['upper','_u','u'] =  30
 
     mpc.setup()
 

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_mujoco/main.py

@@ -25,14 +25,19 @@ from OpenGL.GLU import *
 from OpenGL.GLUT import *
 
 # make function to do co design using finite difference
-def balance(delta_t = 0.02, plotting = False, polelen = .3):
+def balance(delta_t = 0.02, plotting = False, cart_m = 10.5, kg_m_pole = 16.687):
 
     # set initial conditions
     x0 = [0, np.deg2rad(180)]
     model, data = mjmod_init(x0)
+ 
+    # set pole length and update mass/intertia
+    #model.body_mass[2] = polelen*kg_m_pole
+    #model.body_inertia[1,2] = (1/12)*polelen*kg_m_pole*(polelen**2)
+    #model.body_mass[1] = 4.1
 
-    # set pole length
-    setpolelen(model, data, polelen)
+    #setpolelen(model, data, polelen)
+
 
     # init window
     window, camera, scene, context, viewport = mjrend_init(model, data)
@@ -66,13 +71,16 @@ def balance(delta_t = 0.02, plotting = False, polelen = .3):
         video_writer = cv2.VideoWriter('cartpole.mp4', fourcc, 30.0, (width, height))
 
     stop = False
+    Fail = False
     while(stop==False):
         if step != 1:
-            if abs(curx) < xtol and abs(curtheta) < thetatol and abs(curdx) < dxtol and abs(curdtheta) < dthetatol:
+            if abs(curx) < xtol and abs(curtheta) < thetatol and abs(curdx) < dxtol and abs(curdtheta) < dthetatol: # if success
                 stop = True
-            elif glfw.window_should_close(window):
+            elif glfw.window_should_close(window): # if window is closed
                 stop = True
-
+            #elif abs(curdx) < dxtol and abs(curtheta) > thetatol and abs(curx) > xtol:
+               # stop = True
+               # Fail = True
 
 
         if plotting:
@@ -163,6 +171,7 @@ def balance(delta_t = 0.02, plotting = False, polelen = .3):
 
     # close window
     glfw.terminate()
+    print(float(cost))
     if plotting:
         video_writer.release()
 
@@ -175,8 +184,8 @@ def balance(delta_t = 0.02, plotting = False, polelen = .3):
         from animate_cartpole import animate_cartpole
         animate_cartpole(xarr, thetaarr, farr, gif_fps=20, l=1, save_gif=True, name='cartpole_mjpc.gif')
     tf = tarr[-1]
-    return float(cost), tf
[…]

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_mujoco/model.py

@@ -23,7 +23,7 @@ def model_set(A, B):
     model.set_rhs('x', xdot)
 
     # cost (need to change)
-    J = (x[0])**2 + 50*(x[1])**2 + (x[2])**2 + (x[3])**2 
+    J = 0.2*(x[0])**2 + (x[1])**2 + 0.1*(x[2])**2 + 0.1*(x[3])**2 
     model.set_expression(expr_name='cost', expr=J)
     
 

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_dynamics/animate_cartpole.py
whyyyyy

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 import matplotlib.patches as patches
 from matplotlib import animation
 
-def animate_cartpole(x, theta, force, l=1, interval=20, force_scaler=0.2, save_gif=False, gif_fps=20):
+def animate_cartpole(x, theta, force, l=1, interval=20, force_scaler=0.2, save_gif=False, gif_fps=20, name='cartpole.gif'):
     # x: time history of cart location, 1d vector
     # theta: time history of pole angle, 1d vector
     # force: control input force
@@ -58,8 +58,8 @@ def animate_cartpole(x, theta, force, l=1, interval=20, force_scaler=0.2, save_g
     
     anim = animation.FuncAnimation(fig, animate, init_func=init, frames=len(x), interval=interval, repeat=True)
     if save_gif:
-        anim.save("cartpole.gif", writer=animation.PillowWriter(fps=gif_fps))
-    plt.show()
+        anim.save(name, writer=animation.PillowWriter(fps=gif_fps))
+    #plt.show()
 
 
 if __name__ == '__main__':

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_dynamics/controller.py

@@ -14,13 +14,17 @@ def control (model, delta_t):
     mpc = do_mpc.controller.MPC(model)
 
     setup_mpc = {
-    'n_robust': 5,
-    'n_horizon': 40,
-    't_step': delta_t,
-    'state_discretization': 'discrete',
-    'store_full_solution':False,
-    # Use MA27 linear solver in ipopt for faster calculations:
-    #'nlpsol_opts': {'ipopt.linear_solver': 'MA27'}
+        'n_horizon': 200,
+        'n_robust': 0,
+        'open_loop': 0,
+        't_step': delta_t,
+        'state_discretization': 'collocation',
+        'collocation_type': 'radau',
+        'collocation_deg': 3,
+        'collocation_ni': 1,
+        'store_full_solution': True,
+        # Use MA27 linear solver in ipopt for faster calculations:
+        # 'nlpsol_opts': {'ipopt.linear_solver': 'mumps'}
     }
     mpc.settings.supress_ipopt_output()
     mpc.set_param(**setup_mpc)
@@ -31,22 +35,28 @@ def control (model, delta_t):
 
     mpc.set_objective(mterm=mterm, lterm=lterm)
 
-    mpc.set_rterm(u=1e-6) # input penalty
+    mpc.set_rterm(u=0.1) # input penalty
 
-    max_x = np.array([[8], [30], [100], [100]])
-    min_x = np.array([[-8], [-30], [-100], [-100]])
+    max_x = np.array([[3.5], [200], [1000], [1000]])
+    min_x = np.array([[-3.5], [-200], [-1000], [-1000]])
 
     # lower bounds of the states
-    mpc.bounds['lower','_x','x'] = min_x
+    #mpc.bounds['lower','_x','x'] = min_x[0]
+    mpc.bounds['lower','_x','theta'] = min_x[1]
+    mpc.bounds['lower','_x','dx'] = min_x[2]
+    mpc.bounds['lower','_x','dtheta'] = min_x[3]
 
     # upper bounds of the states
-    mpc.bounds['upper','_x','x'] = max_x
+    #mpc.bounds['upper','_x','x'] = max_x[0]
+    mpc.bounds['upper','_x','theta'] = max_x[2]
+    mpc.bounds['upper','_x','dx'] = max_x[2]
+    mpc.bounds['upper','_x','dtheta'] = max_x[3]
 
     # lower bounds of the input
[…]

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_dynamics/main.py

@@ -12,116 +12,52 @@ import do_mpc
 from model import model_set
 from controller import control
 
-def run_onestep(x0, h, delta_t):
-    
-    #set up system and controller
-    model = model_set(x0, h)
-    mpc = control(model, delta_t)
+m = 1
+M = 5
+L = 1
 
-    estimator = do_mpc.estimator.StateFeedback(model)
-    simulator = do_mpc.simulator.Simulator(model)
+num_steps = 200
 
-    simulator.set_param(t_step = delta_t)
-    simulator.setup()
+delta_t = .04
+model = model_set(M,m,L)
+mpc = control(model, delta_t)
 
-    # Initial state
-    mpc.x0 = x0
-    simulator.x0 = x0
-    estimator.x0 = x0
+# estimator and simulator (need to replace with mujoco)
+estimator = do_mpc.estimator.StateFeedback(model)
+simulator = do_mpc.simulator.Simulator(model)
+simulator.set_param(t_step = delta_t)
+simulator.setup()
 
-    # Use initial state to set the initial guess.
-    mpc.set_initial_guess()
+x0 = np.array([0, 0, 0, 0])
+# Initial state
+mpc.x0 = x0
+simulator.x0 = x0
+estimator.x0 = x0
 
-    # preform steps
+# Use initial state to set the initial guess.
+mpc.set_initial_guess()
+u0 = 0
+
+# control   
+xarr = []
+thetaarr = []
+farr = []
+for i in range(num_steps):
     u0 = mpc.make_step(x0)
-    y_next = simulator.make_step(u0)
-    x0 = estimator.make_step(y_next)
+    x0 = simulator.make_step(u0)
+    print(i)
 
-    curx = x0[0]
-    curtheta = x0[1]
-    curdx = x0[2]
-    curdtheta = x0[3]
-    curf = u0[0]
-    J = (curtheta - np.pi)**2
+    curx = float(x0[0])
+    curtheta = float(x0[1])
+    curf = float(u0)
 
-    state = np.matrix([curx, curtheta, curdx, curdtheta])
+    xarr.append(curx)
+    thetaarr.append(curtheta)
+    farr.append(curf)
 
-    return state, curf, J
+from animate_cartpole import animate_cartpole
+animate_cartpole(xarr, thetaarr, farr, gif_fps=20, l=L, save_gif=True, name='cartpole_mjpc.gif')
 
-
-
-
-
[…]

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_dynamics/model.py

@@ -10,28 +10,26 @@ sys.path.append(rel_do_mpc_path)
 # Import do_mpc package:
 import do_mpc
 
-def model_set(X0, h):
-    model_type = 'discrete' # either 'discrete' or 'continuous'
+def model_set(M,m,L):
+    g = 9.81
+    model_type = 'continuous' # either 'discrete' or 'continuous'
     model = do_mpc.model.Model(model_type)
 
-    _x = model.set_variable(var_type='_x', var_name='x', shape=(4,1))
-    _u = model.set_variable(var_type='_u', var_name='u', shape=(1,1))
+    x = model.set_variable(var_type='_x', var_name='x', shape=(1,1))
+    dx = model.set_variable(var_type='_x', var_name='dx', shape=(1,1))
+    theta = model.set_variable(var_type='_x', var_name='theta', shape=(1,1))
+    dtheta = model.set_variable(var_type='_x', var_name='dtheta', shape=(1,1))
+    u = model.set_variable(var_type='_u', var_name='u', shape=(1,1))
 
-    from cartpole_sys import discretize_sys
-    theta = float(X0[1])
-    dtheta = float(X0[3])
-    A, B, C = discretize_sys(theta, dtheta, h)
+    xdd = (-m*g*np.sin(theta)*np.cos(theta) - (u + m*L*(dtheta**2)*np.sin(theta))) / (m*np.cos(theta)**2 - (M+m))
+    thetadd = ((M+m)*g*np.sin(theta) + np.cos(theta)*(u + M*L*(dtheta**2)*np.sin(theta))) / (m*L*np.cos(theta)**2 - (M+m)*L)
 
+    model.set_rhs('x', dx)
+    model.set_rhs('theta', dtheta)
+    model.set_rhs('dtheta', thetadd)
+    model.set_rhs('dx', xdd)
 
-    x_next = A@_x + B@_u + C
-    #print(A)
-    #print(B)
-    #print(C)
-
-
-    model.set_rhs('x', x_next)
-
-    J = 3*(_x[0])**2 + 50*((_x[1] - np.pi)**2) + 10*(_x[2]**2) + (_x[3]**2)
+    J = (x)**2 + 50*(theta - np.pi)**2 + (dx)**2 + (dtheta)**2
     model.set_expression(expr_name='cost', expr=J)
     
 

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_mujoco/controller.py

@@ -15,12 +15,12 @@ def control (model, delta_t):
 
     setup_mpc = {
     'n_robust': 0,
-    'n_horizon': 50,
+    'n_horizon': 200,
     't_step': delta_t,
     'state_discretization': 'collocation',
     'collocation_type': 'radau',
-    'collocation_deg': 2,
-    'collocation_ni': 2,
+    'collocation_deg': 1,
+    'collocation_ni': 1,
     'store_full_solution':False,
     # Use MA27 linear solver in ipopt for faster calculations:
     #'nlpsol_opts': {'ipopt.linear_solver': 'MA27'}
@@ -34,10 +34,10 @@ def control (model, delta_t):
 
     mpc.set_objective(mterm=mterm, lterm=lterm)
 
-    mpc.set_rterm(u=0) # input penalty
+    mpc.set_rterm(u=.001) # input penalty
 
-    max_x = np.array([[1], [100], [100], [100]])
-    min_x = np.array([[-1], [-100], [-100], [-100]])
+    max_x = np.array([[2.5], [200], [1000], [1000]])
+    min_x = np.array([[-2.5], [-200], [-1000], [-1000]])
 
     # lower bounds of the states
     mpc.bounds['lower','_x','x'] = min_x
@@ -46,10 +46,10 @@ def control (model, delta_t):
     mpc.bounds['upper','_x','x'] = max_x
 
     # lower bounds of the input
-    mpc.bounds['lower','_u','u'] = -30
+    mpc.bounds['lower','_u','u'] = -40
 
     # upper bounds of the input
-    mpc.bounds['upper','_u','u'] =  30
+    mpc.bounds['upper','_u','u'] =  40
 
     mpc.setup()
 

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_mujoco/main.py

@@ -4,7 +4,6 @@ import glfw
 # Add do_mpc to path. This is not necessary if it was installed via pip.
 import sys
 import os
-import mujoco_viewer
 rel_do_mpc_path = os.path.join('..','..')
 sys.path.append(rel_do_mpc_path)
 
@@ -25,18 +24,18 @@ from OpenGL.GLU import *
 from OpenGL.GLUT import *
 
 # make function to do co design using finite difference
-def balance(delta_t = 0.02, plotting = False, cart_m = 10.5, kg_m_pole = 16.687):
+def balance(delta_t = 0.02, plotting = False, polelen = 0.3, kg_m_pole = 16.687):
 
     # set initial conditions
     x0 = [0, np.deg2rad(180)]
     model, data = mjmod_init(x0)
  
     # set pole length and update mass/intertia
-    #model.body_mass[2] = polelen*kg_m_pole
-    #model.body_inertia[1,2] = (1/12)*polelen*kg_m_pole*(polelen**2)
-    #model.body_mass[1] = 4.1
+    model.body_mass[2] = polelen*kg_m_pole
+    model.body_inertia[1,2] = (1/12)*polelen*kg_m_pole*(polelen**2)
+    #model.body_mass[1] = 9
 
-    #setpolelen(model, data, polelen)
+    setpolelen(model, data, polelen)
 
 
     # init window
@@ -188,4 +187,4 @@ def balance(delta_t = 0.02, plotting = False, cart_m = 10.5, kg_m_pole = 16.687)
 
 
 if __name__ == "__main__":
-    balance(plotting=True)
+    balance(polelen = 0.4,plotting=True)

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_mujoco/mj_interface.py

@@ -44,7 +44,7 @@ def mjrend_init(model, data):
     scene = mujoco.MjvScene(model, 6000)
     camera = mujoco.MjvCamera()
     camera.trackbodyid = 2
-    camera.distance = 8
+    camera.distance = 10
     camera.azimuth = 90
     camera.elevation = -20
     mujoco.mjv_updateScene(

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_dynamics/controller.py
ytrwe

@@ -35,28 +35,28 @@ def control (model, delta_t):
 
     mpc.set_objective(mterm=mterm, lterm=lterm)
 
-    mpc.set_rterm(u=0.1) # input penalty
+    mpc.set_rterm(u=0.01) # input penalty
 
-    max_x = np.array([[3.5], [200], [1000], [1000]])
-    min_x = np.array([[-3.5], [-200], [-1000], [-1000]])
+    max_x = np.array([[5], [200], [1000], [1000]])
+    min_x = np.array([[-5], [-200], [-1000], [-1000]])
 
     # lower bounds of the states
-    #mpc.bounds['lower','_x','x'] = min_x[0]
+    mpc.bounds['lower','_x','x'] = min_x[0]
     mpc.bounds['lower','_x','theta'] = min_x[1]
     mpc.bounds['lower','_x','dx'] = min_x[2]
     mpc.bounds['lower','_x','dtheta'] = min_x[3]
 
     # upper bounds of the states
-    #mpc.bounds['upper','_x','x'] = max_x[0]
+    mpc.bounds['upper','_x','x'] = max_x[0]
     mpc.bounds['upper','_x','theta'] = max_x[2]
     mpc.bounds['upper','_x','dx'] = max_x[2]
     mpc.bounds['upper','_x','dtheta'] = max_x[3]
 
     # lower bounds of the input
-    mpc.bounds['lower','_u','u'] = -50
+    mpc.bounds['lower','_u','u'] = -25
 
     # upper bounds of the input
-    mpc.bounds['upper','_u','u'] =  50
+    mpc.bounds['upper','_u','u'] =  25
 
     mpc.setup()
 

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_dynamics/main.py

@@ -12,9 +12,9 @@ import do_mpc
 from model import model_set
 from controller import control
 
-m = 1
-M = 5
-L = 1
+m = 0.2
+M = 0.6
+L = 0.5
 
 num_steps = 200
 
@@ -42,21 +42,41 @@ u0 = 0
 xarr = []
 thetaarr = []
 farr = []
+tarr = []
 for i in range(num_steps):
     u0 = mpc.make_step(x0)
     x0 = simulator.make_step(u0)
     print(i)
 
     curx = float(x0[0])
-    curtheta = float(x0[1])
+    curtheta = float(x0[2])
     curf = float(u0)
+    curt = i*delta_t
 
     xarr.append(curx)
     thetaarr.append(curtheta)
     farr.append(curf)
+    tarr.append(curt)
 
 from animate_cartpole import animate_cartpole
 animate_cartpole(xarr, thetaarr, farr, gif_fps=20, l=L, save_gif=True, name='cartpole_mjpc.gif')
+
+import matplotlib.pyplot as plt
+fig, (ax1, ax2, ax3) = plt.subplots(3)
+fig.suptitle('States and Controls Over Entire Range')
+fig.tight_layout()
+
+# position states
+ax1.plot(tarr, xarr)
+ax2.plot(tarr, thetaarr)
+ax3.plot(tarr, farr)
+
+ax1.set_ylabel('X')
+ax2.set_ylabel('Theta')
+ax3.set_ylabel('F')
+
+ax3.set_xlabel('Time')
+plt.savefig('timeseries_dynamics', bbox_inches='tight')
 
 
 

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_dynamics/model.py

@@ -21,15 +21,16 @@ def model_set(M,m,L):
     dtheta = model.set_variable(var_type='_x', var_name='dtheta', shape=(1,1))
     u = model.set_variable(var_type='_u', var_name='u', shape=(1,1))
 
-    xdd = (-m*g*np.sin(theta)*np.cos(theta) - (u + m*L*(dtheta**2)*np.sin(theta))) / (m*np.cos(theta)**2 - (M+m))
-    thetadd = ((M+m)*g*np.sin(theta) + np.cos(theta)*(u + M*L*(dtheta**2)*np.sin(theta))) / (m*L*np.cos(theta)**2 - (M+m)*L)
+    #xdd = (-m*g*np.sin(theta)*np.cos(theta) - (u + m*L*(dtheta**2)*np.sin(theta))) / (m*np.cos(theta)**2 - (M+m))
+    xdd = (u - m*L*np.sin(theta)*(dtheta**2) + m*L*np.cos(theta)*dtheta) / (M + m)
+    thetadd = (m*L*np.cos(theta)*xdd + m*g*L*np.sin(theta)) / (m*L**2)
 
     model.set_rhs('x', dx)
     model.set_rhs('theta', dtheta)
     model.set_rhs('dtheta', thetadd)
     model.set_rhs('dx', xdd)
 
-    J = (x)**2 + 50*(theta - np.pi)**2 + (dx)**2 + (dtheta)**2
+    J = (x)**2 + 50*(theta - np.pi)**2 #+ (dx)**2 + (dtheta)**2
     model.set_expression(expr_name='cost', expr=J)
     
 

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_mujoco/controller.py

@@ -14,16 +14,17 @@ def control (model, delta_t):
     mpc = do_mpc.controller.MPC(model)
 
     setup_mpc = {
-    'n_robust': 0,
-    'n_horizon': 200,
-    't_step': delta_t,
-    'state_discretization': 'collocation',
-    'collocation_type': 'radau',
-    'collocation_deg': 1,
-    'collocation_ni': 1,
-    'store_full_solution':False,
-    # Use MA27 linear solver in ipopt for faster calculations:
-    #'nlpsol_opts': {'ipopt.linear_solver': 'MA27'}
+        'n_horizon': 200,
+        'n_robust': 0,
+        'open_loop': 0,
+        't_step': delta_t,
+        'state_discretization': 'collocation',
+        'collocation_type': 'radau',
+        'collocation_deg': 3,
+        'collocation_ni': 1,
+        'store_full_solution': True,
+        # Use MA27 linear solver in ipopt for faster calculations:
+        'nlpsol_opts': {'ipopt.linear_solver': 'mumps'}
     }
     mpc.settings.supress_ipopt_output()
     mpc.set_param(**setup_mpc)
@@ -34,10 +35,10 @@ def control (model, delta_t):
 
     mpc.set_objective(mterm=mterm, lterm=lterm)
 
-    mpc.set_rterm(u=.001) # input penalty
+    mpc.set_rterm(u=0.001) # input penalty
 
-    max_x = np.array([[2.5], [200], [1000], [1000]])
-    min_x = np.array([[-2.5], [-200], [-1000], [-1000]])
+    max_x = np.array([[3.5], [200], [1000], [1000]])
+    min_x = np.array([[-3.5], [-200], [-1000], [-1000]])
 
     # lower bounds of the states
     mpc.bounds['lower','_x','x'] = min_x
@@ -46,10 +47,10 @@ def control (model, delta_t):
     mpc.bounds['upper','_x','x'] = max_x
 
     # lower bounds of the input
-    mpc.bounds['lower','_u','u'] = -40
+    mpc.bounds['lower','_u','u'] = -35
 
     # upper bounds of the input
-    mpc.bounds['upper','_u','u'] =  40
+    mpc.bounds['upper','_u','u'] =  35
 
     mpc.setup()
 

### 2024-12-18 code/nl_mpc_example/dompc/cartpole_mujoco/main.py
vis

@@ -187,4 +187,4 @@ def balance(delta_t = 0.02, plotting = False, polelen = 0.3, kg_m_pole = 16.687)
 
 
 if __name__ == "__main__":
-    balance(polelen = 0.4,plotting=True)
+    balance(polelen = 0.3,plotting=True)

### 2024-12-18 article_fall24/main.tex
update to journal

--- 2024-05-07 R0_journal/main.tex
@@
 
-\title{Research project title}
+\title{Summary of Fall 24 Research}
 \author{Max Howell}
@@
 \begin{document}
-\bibliographystyle{plain}
-\bibliography{references}
-
 \maketitle{}
-
+\bibliographystyle{ieeetr}
 \begin{abstract}
-
-Control Co-Design (CCD) is a method used to optimally design mechanical systems which involves desinging the systems parameters and controling the system at the same time. This method 
-had potential to be less time consuming and more efficent than traditional methods used to design systems where an iterative process is used. In this paper, CCD is applied to a 
-bipedal walking robot known as the compass gait in order to optimize the masses of the upper body and legs and the length of the legs in order to reduce the torque needed to control 
-the system and reduce the time per cycle when walking. 
-A trajectory generation method was also used in this paper to ensure stability of the compass gait system, as the passive compass gait has a very small range of initial conditions for 
-which the system is stable for.
-During the optimization process the cost of control for the compass gait was reduced by 10.83$\%$
-when compared to a compass gait using only open loop control. The torque needed to control the compass gait was also reduced by 78.2$\%$ and the time per cycle was reduced by 10.6$\%$.
-It was also shown that even a simple optimization problem of the passive compass gait can increase the range of stability for the system. When a passive compass gait was optimized to maximize the hip 
-mass, it was shown that not only can the hip mass be optimized by chaning the system parameters, but when simulating the optimized compass gait with 
-the same initial conditions which cause a unstable passive system, the optimized system was stable.
 
@@
 \section{Introduction}
-% introduce control-codesign
-Control Co-Design (CCD) is a design method used to optimally design a system's parameters and control the system at the same time. Traditionally in engineering,
-when a system is designed an iterative process is used to determine the best parameters for the system; however, this can be time consuming and inefficient. 
-It has been shown in \cite{} that the cost of control can be decreased by atleast xx\% in systems such as a cartpole pendulum and a quadrotor by using CCD. CCD applications range
-from design optimizatoin of (examples of CCD being used with citations)
 
-%introduce dymos a a way to use CCD for walking bipdel system
-CCD can be used in junciton with optimal control algorithms with the objective of minimizing or maximizing a desired cost function. One 
-excellent resource used in this paper to perform both CCD and optimal control in a bipedal walking robot is Dymos \cite{Falck2021}. The Dymos multidisciplinary optimal control library
-provides a platform to build and control systems with a variety of different control algorithms.
-
-%
-In this paper, CCD is used to design the system parameters of an underactuated bipedal walking robot known as the compass gait. The limit cylce
-of the compass gait model aswell as the effect of changing system parameters has been thoroughly studied 
-in \cite{Goswami} and \cite{goswami:inria-00073701}. Control of the compass gait robot as also been acheived using passive methods such as potential energy shaping
-\cite{Spong} and active methods such as MPC Control \cite{Kamath2009}; however, CCD has not been applied to the compass gait model.
-In this paper CCD is applied to the compass gait model using open loop control and trajectory optimization with Dymos.
-
-\section{Introduction to the Compass Gait}
-The compass gait model is an underactuated two legged robot capable of passively walking downhill under the right initial conditions.
-Shown in Figure \ref{fig:compassgaitmodel} is a schematic showing the compass gait and its respective system parameters. The two leg masses are modeled as
-point masses with leg lengths a and b below and above the point mass making a total leg length l. The upper body of the robot is approximated as a point mass
-applied at the hip, $m_H$. The state variables of the system are the swing leg angle (the back leg) $\theta_{sw}$ and the stance leg (the front leg) angle
-$\theta_{st}$. A torque, $\tau$ is applied at the hip to each leg in the opposite direction. The angle of the ramp which the compass gait walks on is $\gamma$.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=10cm]{./figures/compassgaitmodel.png}
-\caption{Schematic of the compass gait model. Angles $\theta_{sw}$ and $\theta_{st}$ are shown as the swing leg angle and stance leg angle, respectively.}
-\label{fig:compassgaitmodel}
-\end{figure}
-
-% assumptions for model
-In order to use the compass gait model a few assumptions must be made about the system. First, since the compass gait does not have the ability to bend at the hip, it must 
-be assumed that during the swing, the swing leg retracts so that it does not collide with the ground during the swing phase. 
-It is also assumed that a perfectly inelastic collision without sliding
-occurs the instant the swing leg hits the ground. The compass gait is an extremely simple model, however it should be noted that physical models of the compass gait
-have been built in \cite{article}.
-
-% intro of eqs
-In order to create a state space realization of the compass gait model a state matrix must be defined. For simplicity the swing and stance angles are renamed to
-$x_1$ and $x_2$, respectively, with $x_3$ and $x_4$ being the anguar velocities of the swing and stance legs. The state matrix $\mb{X}$ is shown in equation
-\ref{eq:statematrix}.
-
-\begin{equation}
-\label{eq:statematrix}
-\mb{X} = 
-\begin{bmatrix}
-\\x_1\\
-\\x_2\\
-\\x_3\\
-\\x_4\\
-\end{bmatrix}, \quad
-\begin{bmatrix}
-\\x_3\\
-\\x_4\\
-\end{bmatrix}
- = 
-\begin{bmatrix}
-\\\dot{x_1}\\
-\\\dot{x_2}\\
-\end{bmatrix}
-\end{equation}
-
-The dynamic equations of the swing stage of the compass gait, from \cite{french confrence}, can be written in standard planar manupular dynamics form. 
-This is shown in \ref{eq:swingeqn}, where $\mb{M}$, $\mb{N}$, and $\mb{G}$ are matrices shown in the appendix, and $\mb{U}$
- is the vector used to apply the torque to each leg, shown in
-\ref{eq:U}.
-
-\begin{equation}
-\label{eq:swingeqn}
-\mb{M(X)}\mb{\ddot{X}} + \mb{N(X, \dot{X})}\mb{\dot{X}} + \mb{G(X)} = \tau\mb{U}
-\end{equation}
-
-\begin{equation}
-\label{eq:U}
-\mb{U} = 
-\begin{bmatrix}
-\\1\\
-\\-1\\
-\\0\\
-\\0\\
-\end{bmatrix}
-\end{equation}
-
-Since heel strike (when the swing leg collides with the ground) is modeled as a perfectly inelastic 
-collision, there is an instantaneous angular velocity change when heel stike occurs. Also, since the swing leg becomes the stance leg
-after heel strike, the angles $x_1$ and $x_2$ must be switched in the mathematical model after each walking cycle. The transition matrices used to switch $x_1$ and $x_2$ as well 
-as calculate the updated angular velocities after heel strike are shown in \ref{eq:trans1} and \ref{eq:trans2}, where $x^+$ represents the state
-variable after heel strike and $x^-$ represents the state variable before impact, $\mb{Q^-}$ and $\mb{Q^+}$ are the angular velocity transition matrices shown in the appendix.
-
-\begin{equation}
-\label{eq:trans1}
-\begin{bmatrix}
-\\x_1^+\\
-\\x_2^+\\
-\end{bmatrix}
- = 
-\begin{bmatrix}
-0 & 1\\
-1 & 0\\
-\end{bmatrix}
-\begin{bmatrix}
-\\x_1^-\\
-\\x_2^-\\
-\end{bmatrix}
-\end{equation}
-
-\begin{equation}
-\label{eq:trans2}
-\mb{Q^+}
-\begin{bmatrix}
-\\x_3^+\\
-\\x_4^+\\
-\end{bmatrix}
- = 
-\mb{Q^-}
-\begin{bmatrix}
-\\x_3^-\\
-\\x_4^-\\
-\end{bmatrix}
-\end{equation}
-
-\section{Natural Limit Cycles of the Passive Compass Gait}
-
-It is important to define stability in the context of the compass gait for this paper. One possible meaning of stability in this context is that, while walking, 
-the compass gait procedes forwards in a controlled manner and takes the same "path" for each walking cycle. Or perhaps stability could be mathematically defined
-using Poincaré maps. While both of these definitions of stability are valid, stability in this paper simply refers to the ability of the compass gait system to 
-continue walking forwards for the defined number of cycles.
-
-When simulating the passive compass gait system, it can be observed that the stability of the compass gait is heavily dependent on the
-given initial conditions and system parameters. It can also be observed that the basin of stability is extremely small and it is almost impossible
-to "guess" the correct IC's and system parameters to create a stable walking system. This is one of the main issues of the compass gait.
-
-When using stable initial conditions, the compass gait can be simulated over one or multiple walking cycles. Shown in Figure \ref{fig:onecycle}
-is an example of a stable limit cycle for the compass gait over one walking cycle. As can be seen, the swing leg makes a large arc during the swing phase of the 
-compass gait, while the stance leg pivots in place. The state variables during this walking cycle are also shown over time in Figure \ref{fig:states_onecycle}.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=10cm]{./figures/onecycle.pdf}
-\caption{One walking limit cycle of the compass gait. The swing leg makes a large swing (shown in blue, a large cicle) around while the stance leg only pivots (shown in orange.)}
-\label{fig:onecycle}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=9cm]{./figures/states_onecycle.pdf}
-\caption{States of the compas gait over time for one walking cycle.}
-\label{fig:states_onecycle}
-\end{figure}
-
-Figures \ref{fig:onecycle} and \ref{fig:states_onecycle} show the compass gait when stable initial conditions are used; however, most initial conditions
-given to the compass gait system result in an unstable system. An example of a unstable gait is shown in Figure \ref{fig:unstable_onecycle}, where the initial angular velocities are 
-not sufficent to overcome gravity, and the compass gait cannot start a cycle and will eventually fall down. It can be seen in this figure, that the swing leg initial angle 
-is approximataly equal to the final angle of the swing leg, and the stance leg starts going forward, but runs out of momentum before falling backwards.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=10cm]{./figures/unstable_onecycle.pdf}
-\caption{An example of an unstable limit cycle. In this case, the initial angular velocities make it impossible for the compass gait to start a walking cycle.}
-\label{fig:unstable_onecycle}
-\end{figure}
-
-\clearpage
-
-\section{Passive Optimization with Co-Design}
-
-In order to acheive co-design with a simple passive compass gait system, an optimization problem was proposed.
-This optimization problem, outlined in Figure \ref{fig:optsetup}, involved maximizing the point mass at the hip ($m_H$) in order to simulate the
-compass gait robot "carrying" a heavy object. In order to acheive this goal, the lengths below and above the leg point mass, a and b, were optimized.
-The leg point mass ($m$) also changed based on the total length of the legs based on the given density of the legs.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=6cm]{./figures/opt1prob.png}
-\caption{Outline of the simple passive optimization problem. The goal is to incrase the range of stable initial conditions while maximizing the applied hip mass.}
-\label{fig:optsetup}
-\end{figure}
-
-The goal of this optimization problem was to increase the range of initial conditions that caused the compass gait to complete one stable cycle
-while maximizing the mass at the hip. In order to acheive this goal a trajectory generation method was used, where the compass gait model was
-first simulated using known stable initial conditions and system parameters and the final position states ($x_1$ and $x_2$) were recorded and passed to the optimization
-model. The final position states of the optimization model were constrained to be equal to the recorded final position states found in the trajectory generation
-simulation. This gave the the compass gait optimization problem a goal to acheive in order to keep the system stable. Figure \ref{fig:optlogicdiagram} shows a diagram
-outlining this process.
-
-\begin{figure}[h]
-\setlength{\unitlength}{0.14in} % selecting unit length
-\centering % used for centering Figure
-\begin{picture}(32,15) % picture environment with the size (dimensions)
-% 32 length units wide, and 15 units high.
-\put(0,5.5){\vector(1,0){5}}
-\put(5,3.25){\framebox(10,5){Trajectory Generation}}
-\put(20.2,3.25){\framebox(10,5){Optimization Problem}}
-\put(15,5.5){\vector(1,0){5.2}}
-\put(30.2,5.5){\vector(1,0){12}}
-\put(25.2, 12){\vector(0,-1){3.7}}
-\put(0,6){Stable IC's}
-\put(15.2, 6){End States}
-\put(15.3, 4.5){$x_1$ and $x_2$}
-\put(30.4, 6){Stable, optimized system}
-\put(23.3, 12.2){New IC's}
-\end{picture}
-\caption{A trajectory generation method was used where a system with stable initial conditions found end states ($x_1$ and $x_2$) to be used for the optimization problem.} % title of the Figure
-\label{fig:optlogicdiagram} % label to refer figure in text
-\end{figure}
-
-While it was not possible to make all initial conditions stable without a controller, the range of initial conditions was
-increased. Shown in Figure \ref{fig:nonoptlimcycle} is the limit cycle of the compass gait without optimization with the initial conditions and system parameters 
-shown in Table \ref{tab:nonoptparam}. As can be seen, the compass gait system is not able to complete a full cycle and can be said to be unstable.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=10cm]{./figures/nonopt_limitcycle.pdf}
-\caption{Limit cycle of the non optimized compass gait over one cycle with the same initial conditions as the optimized system. It can be seen that the system does not 
-complete a full cycle.}
-\label{fig:nonoptlimcycle}
-\end{figure}
-
-\begin{table}[h]
-\centering
-\caption{Initial Conditions and System Parameters of the Non-Optimized Compass Gait Shown in \ref{fig:nonoptlimcycle}}
-\begin{tabular}{lr}
-\toprule
-Parameter & Value \\
-\midrule
-$m (kg)$ & 5 \\
-$m_H (kg)$ & 10 \\
-$a (m)$ & 0.5 \\
-$b (m)$ & 0.5 \\
-$\gamma (slope, rad)$ & 0.05 \\
-$x_1 (rad)$ & -0.23 \\
-$x_2 (rad)$ & 0.4 \\
-$x_3 (rad/s)$ & -0.24 \\
-$x_4 (rad/s)$ & -1.1 \\
-\bottomrule
-\end{tabular}
-\label{tab:nonoptparam}
-\end{table}
-
-The limit cycle of the non-optimized system can be compared to that of the optimized system in order to see the benefits of the optimization.
-Shown in Figure \ref{fig:passiveopt} is the limit cycle of the optimized compass gait system with the same initial conditions as the non-optimized compass gait.
-It can be seen that after optimization, the system completes a full stable walking cycle.
-The optimized system parameters of the new system are shown in Table \ref{tab:optparam}. Not only can it be seen that the compass gait system gained stability from the 
-optimization, but it also succesfully maximized the mass at the hip.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=10cm]{./figures/passiveopt_limitcycle.png}
-\caption{The limit cycle of the optimized compass gait. It can be seen that the system is stable for the first walking cycle.}
-\label{fig:passiveopt}
-\end{figure}
-
-\begin{table}[h]
-\centering
-\caption{Initial Conditions and System Parameters of the Non-Optimized Compass Gait Shown in \ref{fig:nonoptlimcycle}}
-\begin{tabular}{lr}
-\toprule
-Parameter & Value \\
-\midrule
-$m (kg)$ & 8.96 \\
-$m_H (kg)$ & 80 \\
-$a (m)$ & 0.83 \\
-$b (m)$ & 0.957 \\
-\bottomrule
-\end{tabular}
-\label{tab:optparam}
-\end{table}
-
-While this optimization problem was succesful, it has its limitations. First, as previously mentioned, without a using an active controller
-it is impossible to guarentee that all initial conditions for the compass gait will result in a stable walking limit cycle. Also, it is desired that the optimized
-compass gait system complete multiple walking cycles instead of just one. In order to fix these two limitation, active CCD can be used to optimize the problem 
-and simulate the compass gait over any given number of walking cycles.
-
-\clearpage
-
-\section{Active Control Co-Design of the Compass Gait}
-
-In order to perform Active CCD on the compass gait a torque controller was activated at the hip. The addition of this torque controller allowed
-for the problem to be optimized to a more meaningful cost function. Instead of simply optimizing the hip mass of the compass gait, the time per cycle 
-and the total torque used was minimized. The cost function used is shown in \ref{eq:costfunc}, where $\tau_r$ and $t_r$ are reference scalers used 
-to normalize $\tau$ and $t$. The values of $\tau_r$ and $t_r$ which were used are shown in \ref{eq:refvars}
-
-\begin{equation}
-\label{eq:costfunc}
-J = 
-\int_{0}^{t} (\tau / \tau_r)^2 + (t/t_r) \,dt 
-\end{equation}
-
-\begin{equation}
-\label{eq:refvars}
-\tau_r = 6.32 (N*m), 
-t_r = 0.5 (s)
-\end{equation}
-
-Since the hip mass was no longer part of the cost function, it was chosen to also 
-be a design variable of the system. The constraints of the system remained the same, outlined in Figure \ref{fig:OLprob}.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=7cm]{./figures/openloopcontrol.png}
-\caption{The torque and total time to complete a cycle were minimized in the active open loop CCD problem.}
-\label{fig:OLprob}
-\end{figure}
-
-In order to optimize the compass gait through multiple walking cycles and ensure stability, the trajectory generation method used
-in the previous section was used with continued simulation of the system after the first cycle.
-After the trajectory generation was performed and the final states of the system were determined, CCD was performed on the system in order to
-optimize the design variables and the torque over one walking cycle. After the optimal values for the design variables were determined, the system was 
-simulated for a fixed number of cycles with trajectory optimization (torque optimization) according to the set cost function. Figure \ref{fig:OLblock}
-shows a diagram outlining the new trajectory optimization process.
-
-\begin{figure}[h]
-\setlength{\unitlength}{0.14in} % selecting unit length
-\centering % used for centering Figure
-\begin{picture}(32,15) % picture environment with the size (dimensions)
-% 32 length units wide, and 15 units high.
-\put(0,5.5){\vector(1,0){5}}
-\put(5,3.25){\framebox(10,5){Trajectory Generation}}
-\put(20.2,3.25){\framebox(8,5){CCD For 1 Cycle}}
-\put(30.2,3.25){\framebox(12,5){Simulation with Torue Opt.}}
-\put(15,5.5){\vector(1,0){5.2}}
-\put(28.2,5.5){\vector(1,0){2}}
-\put(25.2, 12){\vector(0,-1){3.7}}
-\put(10, 3.25){\vector(0,-1){2}}
-\put(10, 1.25){\vector(1, 0){26.2}}
-\put(36.2, 1.25){\vector(0, 1){2}}
-\put(0,6){Stable IC's}
-\put(15.2, 6){End States}
-\put(15.3, 4.5){$x_1$ and $x_2$}
-\put(23.3, 12.2){New IC's}
-\put(18, 1.5){End States $x_1$ and $x_2$}
-\end{picture}
-\caption{The trajectory generation method used in the previous example was used with continued trajectory optimization after one cycle.} % title of the Figure
-\label{fig:OLblock} % label to refer figure in text
-\end{figure}
-
-In order to compare the compass gaits performance when CCD is used to other compass gait scenarios, three simulations were ran and compared:
-The comass gait with active CCD (Figure \ref{fig:OLblock}), the compass gait with active open loop control without co-design, and the non-optimized passive compass gait system.
-Each of these systems were simulated over 10 cycles and the stability, total torque used, time per cycle, and optimized design variables were compared. Each system used the initial conditions and 
-system parameters shown in Table \ref{tab:params}.
-
-\begin{table}[h]
-\centering
-\caption{Initial Conditions and System Parameters Used to Simulate the Compass Gait.}
-\begin{tabular}{lr}
-\toprule
-Parameter & Value \\
-\midrule
-$m (kg)$ & 5 \\
-$m_H (kg)$ & 10 \\
-$a (m)$ & 0.5 \\
-$b (m)$ & 0.5 \\
-$\gamma (slope, rad)$ & 0.05 \\
-$x_1 (rad)$ & -0.3 \\
-$x_2 (rad)$ & 0.2038 \\
-$x_3 (rad/s)$ & -0.41215 \\
-$x_4 (rad/s)$ & -1.0501 \\
-\end{tabular}
-\label{tab:params}
-\end{table}
-
-Shown in Figures \ref{fig:limcyc_passive} and \ref{fig:timehis_passive} are the limit cycle and total time history of the state variables for the passive compass gait system when simulated using the 
-given initial conditions and system parameters. As can be seen the chosen initial conditions resulted in a stable, well behaved system for the passive compass gait. It can be 
-seen that it takes about three cycles for the system to converge to a single limit cycle; however, there still exists some oscilitations in the limit cycle.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=8cm]{./figures/limitcycle_passivesim.pdf}
-\caption{The limit cycle of the passive compass gait over 10 cycles.}
-\label{fig:limcyc_passive}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=8cm]{./figures/timehis_passivesim.png}
-\caption{The total time history of the state variables of the passive compass gait over 10 cycles.}
-\label{fig:timehis_passive}
-\end{figure}
-
-Similarly, shown in Figures \ref{fig:lim_noco} and \ref{fig:timehis_noco} are the limit cycle and total time history of the state variables and controls for the compass gait 
-using active open loop control without co-design. The same trajectory generation method was used during this simulation, except without co-design (only torque optimization).
-It can be seen that it takes approximataly 6 cycles to converge to a single limit cycle; also, the indivual walking cycles are much more spaced out when compared to the passive system.
-However, the final limit cycle is much smaller than that of the passive limit cycle, resulting in less time per cycle. This can be seen when comparing the total time history of the 
-active system without co-design to the passive system; the average time per cycle and total time of the active system is smaller than that of the passive system.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=8cm]{./figures/limcycle_nocodesign.png}
-\caption{The limit cycle of the active compass gait using open loop control without co-design.}
-\label{fig:lim_noco}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=8cm]{./figures/timehis_nocodesign.png}
-\caption{The total time history of the state variables, control, and cost of the active compass gait without co-design.}
-\label{fig:timehis_noco}
-\end{figure}
-
-Finally, the limit cycle and total time history of the compass gait when CCD is used is shown in Figures \ref{fig:limcyc_CCD} and \ref{fig:timehis_CCD}. When comparing the limit 
-cycle of the CCD compass gait to the other two systems, it can be seen that one small waliking cycle is performed before converging into the final limit cycle in approximataly 4 
-iterations. It can also be seen from the time history that the CCD compass gait has a smaller time per cycle than both the passive system and the open loop system without co-design. 
-It can also be seen that much less torque is used to control the system than the compass gait without co-design.
-
-Shown in Table \ref{tab:opt_vars} are the optimized design variables found using CCD with the compass gait system. As can be seen, in order to reduce torque and time per cycle, 
-the system prefers to have a higher mass at the hip and a longer leg segment below the point mass on the leg. In order to determine the best design variables for a system, 
-the bounds of the design variables should be adjusted as desired.
-
-\begin{table}[h]
-\centering
-\caption{Optimized Design Variables Found Using CCD}
-\begin{tabular}{lr}
-\toprule
-Parameter & Value \\
-\midrule
-$m (kg)$ & 6.45 \\
-$m_H (kg)$ & 100 \\
-$a (m)$ & 1 \\
-$b (m)$ & 0.29 \\
-\end{tabular}
-\label{tab:opt_vars}
-\end{table}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=8cm]{./figures/limcyc_CCD.png}
-\caption{The limit cycle of the compass gait when CCD is used.}
-\label{fig:limcyc_CCD}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=8cm]{./figures/timehis_CCD.png}
-\caption{The total time history of the state variables, control, and cost when CCD is used.}
-\label{fig:timehis_CCD}
-\end{figure}
-
-Shown in Table \ref{tab:time} is the average time per cycle and the total time to complete 10 cycles for each compass gait system which was simulated. As can be seen, the CCD system 
-has the shortest average time per cycle and total time, followed by the open loop system without control, then the passive system. From these results it can be said that 
-CCD is more effective in reducing the time per cycle of the compass gait than control without co-design, and any control at all is much better than a passive system.
-
-\begin{table}[h]
-\centering
-\caption{Average Time per Cycle and Total Time for 10 Cycles of Each Simulation (seconds)}
-\begin{tabular}{ccc}
-\toprule
-Simulation & Average Time/Cycle & Total Time \\
-\midrule
-Passive System & 0.680 & 6.8 \\
-Active System Without Co-Design & 0.594 & 5.94 \\
-System with CCD & 0.531 & 5.31\\
-\end{tabular}
-\label{tab:time}
-\end{table}
-
-The total torque used during the simulations of the compass gait with CCD and the open loop system without co-design can also be compared. Shown in Table \ref{tab:tautot} is 
-the summation of the square of $\tau$ over all 10 cycles, this is shown in equation \ref{eq:summation}. From the table, It can be seen that much more torque is needed to control 
-the non-optimized 
-compass gait when compared to the compass gait with CCD. From the results found in these simulations, it can be concluded that using CCD with the compass gait 
-system can greatly decrease the cost of control when compared to compass gait systems without co-design or optimization.
-
-\begin{equation}
-\label{eq:summation}
-\tau_{sum} = 
-\sum_{t = 0}^{t = t_f}
-\tau^2
-\end{equation}
-
-\begin{table}[h]
-\centering
-\caption{Summation of $\tau^2$ for the CCD and Open Loop Control System}
-\begin{tabular}{lr}
-\toprule
-System & $\sum \tau^2$ \\
-\midrule
-Active System Without Co-Design & 133.46\\
-System with CCD & 29.17\\
-\end{tabular}
-\label{tab:tautot}
-\end{table}
-
-\section{Conclusions and Future Works}
-
-During this study, CCD was applied to the compass gait bipedal robot and the system was optimized first to maximize the mass at the hip with a passive system, 
-then to minimize the torque used and time per cycle with a torque controller at the hip. A trajectory generation method was also used in this paper, 
-where a passive system is first simulated with stable initial conditions and the end states ($x_1$ and $x_2$) are used during the optimization process in order to 
-ensure a stable system.
-
-From the passive optimization, it was shown that not only can optimization 
-of the design variables in the compass gait allow for the hip mass to be maximized, but the optimization can also allow for a greater range of initial conditions 
-which result in a stable first walking cycle. It was shown that when the optimized compass gait was given the same initial conditions which would typically 
-result in an unstable walking cycle with a non-optimized passive system, the system parameters were optimized so that the compass gait had a stable first walking cycle 
-while still maximizing the hip mass.
-
-When CCD was used on the active compass gait, it was shown that the cost of control was reduced by 10.83$\%$ when compared to a system 
-with open loop control without co-design using the same intial conditions and system parameters. Shown in Table \ref{tab:costofcon} is the total cost of control over 10 
-walking cycles when CCD is used compared to active control without co-design. The cost shown is the same cost function shown above in equation \ref{eq:costfunc}. CCD reduced the 
-cost by reducing the torque used and time per cycle in accordance with the cost function. The compass gait with CCD used only 21.8$\%$ as much torque as the system without co-design 
-and the time per cycle was reduced by 0.063 seconds, or 10.6$\%$.
-
-\begin{table}[h]
-\centering
-\caption{Cost of Control for the CCD Compass Gait vs Only Open Loop Control}
-\begin{tabular}{lr}
-\toprule
-System & Cost of Control \\
-\midrule
-Active System Without Co-Design & 11.91\\
-System with CCD & 10.62\\
-Percent Improvement & 10.83$\%$ \\
-\end{tabular}
-\label{tab:costofcon}
-\end{table}
-
-One of the main limitations of this study is that CCD was applied to the compass gait over only one walking cycle using the trajectory generation method,
-then the compass gait was simulated with open-loop 
-control for the given number of cycles.
-Ideally, when using CCD for a system with cyclic behavior it would be desired that the system parameters be optimized based on the entire trajectory of the system 
-(over all cycles); however, due to the structure of the Dymos code, it was only possible to optimize the first walking cycle of the compass gait during this study. 
-In future studies, the compass gait should be optimized over the entire trajectory in order to obtain the best possible system parameters for the system.
-
-Another future study which should be conducted is the use of a closed loop control algorithm in order used to perform CCD. One example of this is using 
-MPC or LQR control in order to optimize and control the compass gait. The use of closed loop CCD with the compass would result in more optimal system parameters 
-and better trajectory optimization.
-
-\clearpage
-\appendix
-
-\section{Compass Gait Equations}
-
-Shown in this appendix is a more complete version of the dynamics equations of the compass gait. The swing stage dynamics can be written in standard planar manipulator 
-dynamics form, shown in \ref{eq:swingeq}. The $\mb{M}$, $\mb{N}$, $\mb{G}$, and $\mb{U}$ matrices are also shown in \ref{eq:M}, \ref{eq:N}, \ref{eq:G}, and \ref{eq:U} respectively.
-
-\begin{equation}
-\label{eq:swingeq}
-\mb{M(X)}\mb{\ddot{X}} + \mb{N(X, \dot{X})}\mb{\dot{X}} + \mb{G(X)} = \tau\mb{U}
-\end{equation}
-
-\begin{equation}
-\mb{M} = 
-\label{eq:M}
-\begin{bmatrix}
-\\mb^2 &  -mlbcos(x_2 - x_1)\\
-\\-mlbcos(x_2 - x_1) &  m_{H}l^2 + m(l^2 + a^2)\\
-\end{bmatrix}
-\end{equation}
-
-\begin{equation}
-\mb{N} = 
-\label{eq:N}
-\begin{bmatrix}
-\\0 &  mlb\dot{x_2}sin(x_2 - x_1)\\
-\\-mlb\dot{x_1}sin(x_2 - x_1) &  0\\
-\end{bmatrix}
-\end{equation}
-
-\begin{equation}
-\mb{G} = 
-\label{eq:G}
-\begin{bmatrix}
-\\mgbsin(x_1)\\
-\\-(m_{H}l + m(a+l))gsin(x_2)\\
-\end{bmatrix}
-\end{equation}
-
-\begin{equation}
-\mb{U} = 
-\label{eq:U}
-\begin{bmatrix}
-\\1\\
-\\-1\\
-\\0\\
-\\0\\
-\end{bmatrix}
-\end{equation}
-
-The transition equations when heel strike occurs are shown in \ref{eq:transeqn} and \ref{eq:transeqn2}, where $x^+$ signifies a state variables after heel strike and $x^-$ 
-signals a state variable before
-heel strike. The transition matrices $\mb{Q^+}$ and $\mb{Q^-}$ are shown in \ref{eq:Q-} and \ref{}. Note that $\alpha$ represents half the angle between the swing and stance 
-legs.
-
-\begin{equation}
-\label{eq:transeqn}
-\begin{bmatrix}
-\\x_1^+\\
-\\x_2^+\\
-\end{bmatrix}
- = 
-\begin{bmatrix}
-0 & 1\\
-1 & 0\\
-\end{bmatrix}
-\begin{bmatrix}
-\\x_1^-\\
-\\x_2^-\\
-\end{bmatrix}
-\end{equation}
-
-\begin{equation}
-\label{eq:transeqn2}
-\mb{Q^+}
-\begin{bmatrix}
-\\x_3^+\\
-\\x_4^+\\
-\end{bmatrix}
-= 
-\mb{Q^-}
-\begin{bmatrix}
-\\x_3^-\\
-\\x_4^-\\
-\end{bmatrix}
-\end{equation}
-
-\begin{equation}
-\mb{Q^-} = 
-\label{eq:Q-}
-\begin{bmatrix}
-\\-mab & -mab + (m_Hl^2+2mal)cos(2\alpha)\\
-\\0 & -mab\\
-\end{bmatrix}
-\end{equation}
-
-\begin{equation}
-\mb{Q^+} = 
-\label{eq:Q+}
-\begin{bmatrix}
-\\mb(b-lcos(2\alpha)) & ml(l-bcos(2\alpha)) + ma^2 + m_Hl^2\\
-\\mb^2 & -mblcos(2\alpha)\\
-\end{bmatrix}
-\end{equation}
+\Section{}
 

### 2025-01-07 article_fall24/main.tex
article update

@@
 
+fdas
+
 \end{abstract}
@@
 
-\Section{}
+Control Co-Design (CCD) is an engineering design strategy used to optimize the design of a system with control in mind. CCD can be acheived using various control schemes, 
+including open loop control [1], PID control, and other optimal control methods such as LQR or MPC Control. CCD has a very wide range of applications such as in 
+aerospace systems [2] and in thermal energy storage [3]. While CCD has been used in various industries there has yet to be a compreherensive tool developed
+to preform robust CCD using differentiable optimal control strategies such as nonlinear MPC control.
+
+This research focuses on developing a tool which can be used in order to preform robust CCD using MPC control. This research is primarly focused on robotic applications and uses 
+the MuJoCo [4] physics simulation engine in order to simulate the robotic systems. In this paper, do-mpc [5], a library for nonlinear MPC control, is used to simplify the 
+implementation of MPC control. This research builds off of previous work [6] which used the Dymos Optimal Control library [7] in order to preform CCD of a simple 
+two legged walking robot known as the compass gait.
+
+This portion of research can be broken down into three phases; phase one involved using the work done in [6] and implementing MPC control on both the two legged walker 
+as well as a slightly more complicated three legged walker. The purpose of this phase was to understand how to implement MPC control on a simplified model of a robotic system.
+Phase two of this research consisted of using the knowledge of MPC control from phase one and implementing it in a more complex system using the MuJoCo physics engine.
+Phase three consists of using the MPC controlled MuJoCo system and preforming CCD on the system as well as implementing more complex algorithms used in robotics such as 
+path planning and obstacle avoidance. Phase three of this research is incomplete and is only discussed as a future work in this paper.
+
+\section{Methods}
+
+\subsection{Phase I: Compass Gait and Three Legged Walker}
+
+Phase I consisted of building on the research done in [6] and implementing MPC control on the compass gait system as well as a slightly more complex three legged walker.
+The compass gait is a simple two legged walking robot which is able to passively walk down an incline with no control input. The three legged walker is a similar system 
+but is more realistic as it uses a knee joint to walk. The compass gait as well as three legged walker are hybrid systems, meaning they change dynamics based on the state of the system.
+The compass gait is shown in Figure \ref{fig:compass_gait}; for a more detailed look at the dynamics of the compass gait, see [6].
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=8cm]{./figures/compassgaitmodel.png}
+  \caption{Schematic of the compass gait model. Angles $\theta_{sw}$ and $\theta_{st}$ are shown as the swing leg angle and stance leg angle, respectively.}
+  \label{fig:compass_gait}
+\end{figure}
+
+The three legged walker system is shown in Figure \ref{fig:threeleg}. For a more detailed look at the dynamics of the three legged walker, see [8]. In both the three legged walker and
+compass gait, there is a single control input applied to the hip.
+
+In order to implement MPC control on the compass gait, do-mpc was used with the optimization problem as shown in \ref{eq:twolegopt} where the states of the system ($\mb{x}$)
+are shown in \ref{eq:twolegstates}. In the objective function $x_N$ and $x_k$ represent the terminal and the current state of the system, respectively.
+$x_r$ represents the target state of the system, which is determined to be the state at which both the swing and stance legs are in contact with the ground. This state 
+is shown in Table \ref{tab:twolegterm}.
+
+\begin{equation}
+  \begin{aligned}
+  &\text{Objective: $arg min_{x_k,u_k} (x_N - x_r)^TQ_N(x_N - x_r) + \sum_{k = 0}^{N-1} (x_k - x_r)^TQ(x_k - x_r) + u_k^TRu_k $}\\
+  &\text{Subject to:}\\
+  &\text{Dynamic Equations of System}\\
+  &\text{State Constraints: $-90^{\circ} < x_1 < 90^{\circ }, -90^{\circ} < x_2 < 90^{\circ }  $}
+  \end{aligned}
+  \label{eq:twolegopt}
+\end{equation}
+
+\begin{equation}
+  \mb{x} = 
+  \begin{bmatrix}
+  \\x_{sw}\\
+  \\x_{st}\\
+  \\\dot{x}_{sw}\\
+  \\\dot{x}_{st}\\
+  \end{bmatrix}, \quad
+  \label{eq:twolegstates}
+\end{equation}
+
+\begin{table}[h]
+  \centering
+  \caption{Target states during MPC control of the compass gait.}
+  \begin{tabular}{lr}
+  \toprule
+  State & Target Value \\
+  \midrule
+  $x_{sw}$ & 0.19 \\
+  $x_{st}$ & -0.3 \\
+  \end{tabular}
+  \label{tab:twolegterm}
+\end{table}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=8cm]{./figures/threeleg.png}
+  \caption{Diagram of the three legged walker model where q1, q2, and q3 are the stance, swing, and knee joint angles.}
+  \label{fig:threeleg}
+\end{figure}
+
+The $Q_N$ and $Q$ matrices are the weighting matrices for the MPC control and are simply set to be equal to the identity matrix in this case. 
+The $R$ matrix is the weighting matrix for the control input and is set to be equal to 10 to penalize the single control input.
+
+MPC control was implemented on the three legged walker in a similar manner to the compass gait, except there were two phases considered for this system, a two link and a 
+three link phase. A diagram showing these phases is shown in Figure \ref{fig:threelegstates}. These phases were treated as seperate systems with different governing equations and 
+each phase was controlled using a seperate MPC controller. The transition between the two phases was triggered by "knee-strike", where the 
+knee joint angle became equal to the angle of the swing leg, and "heel-strike", where the swing leg comes into contact with the ground. The 
+transition between the phases were calculated using the transition equations from [8].
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=8cm]{./figures/threelegstates.png}
+  \caption{Diagram showing the two link and three link phases of the three legged walker.}
+  \label{fig:threelegstates}
+\end{figure}
+
+The optimization problem for the three link phase is shown in \ref{eq:threelinkphaseopt} where the states of the system ($x$) consist of 
+the stance, swing, and knee joint angles as well as the angular velocities of these joints. The optimization problem for the the two link phase is 
+shown in \ref{eq:twolinkphaseopt} where the states of the system ($x$) consist of the stance and swing leg angles as well as the angular velocities of these joints.
+
+\begin{equation}
+  \begin{aligned}
+  &\text{Objective: $arg min_{x_k,u_k} (x_N - x_r)^TQ_N(x_N - x_r) + \sum_{k = 0}^{N-1} (x_k - x_r)^TQ(x_k - x_r) + u_k^TRu_k $}\\
+  &\text{Subject to:}\\
+  &\text{Control Range: $-3 < tau < 3$}\\
+  &\text{Dynamic Equations of Three Link Phase}\\
+  &\text{State Constraints: $x_2 > x_3 $}
+  \end{aligned}
+  \label{eq:threelinkphaseopt}
+\end{equation}
+
+\begin{equation}
+  \begin{aligned}
+  &\text{Objective: $arg min_{x_k,u_k} (x_N - x_r)^TQ_N(x_N - x_r) + \sum_{k = 0}^{N-1} (x_k - x_r)^TQ(x_k - x_r) + u_k^TRu_k $}\\
+  &\text{Subject to:}\\
+  &\text{Control Range: $-3 < tau < 3$}\\
+  &\text{Dynamic Equations of Two Link Phase}\\
+  &\text{State Constraints: $x_2 = x_3, \dot{x_2} = \dot{x_3}$}
+  \end{aligned}
+  \label{eq:twolinkphaseopt}
+\end{equation}
+
+Similarly to the compass gait control problem, the $Q_N$ and $Q$ matrices are the identity matrix and the $R$ matrix is set to 1 in this case.
+The target states for each phase are the states which were pre-determined to be the knee strike and heel strike states and are shown in Table \ref{tab:threelegterm}.
+
+\begin{table}[h]
+  \centering
+  \caption{Target states during MPC control of the three legged walker.}
+  \begin{tabular}{lrr}
+  \toprule
+  State & Knee Strike Target & Heel Strike Target\\
+  \midrule
+  $x_{sw}$ & -.106  & -0.29\\
+  $x_{st}$ & 0.326 & 0.19\\
+  $x_{knee}$ & 0.326 & -0.19\\
+  \end{tabular}
+  \label{tab:threelegterm}
+\end{table}
+
+\subsection{Phase II: MuJoCo MPC}
+Phase II of this research consisted of using the knowledge of MPC which was gained from phase I and applying it to a more complex system in the MuJoCo physics engine which can 
+simulate complicated robotic systems such as the humanoid shown in \ref{fig:humanoid}. Instead of applying MPC control to the humanoid, MPC control was first applied to a simpler system 
+in a way which can be easily scaled up to the humanoid. The chosen system was the classic cart pole control problem, where a pole with a 
+mass attatched to the end is attatched to a cart which can move back and forth in one direction. This is shown in \ref{fig:cartpole}. The goal of this problem is to 
+control the pole to balance upright on the cart when the cart is at a stationary set location. This is done by applying a force to the cart to move it in the positive or 
+negative x direction.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=8cm]{./figures/humanoid.png}
+  \caption{Example of a humanoid robot which can be simulated in the MuJoCo physics engine.}
+  \label{fig:humanoid} 
+\end{figure}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=8cm]{./figures/cartpole-dynamics.png}
+  \caption{Diagram of the cart pole system.}
+  \label{fig:cartpole} 
+\end{figure}
+
+The cart pole system has previously been solved with various control techniques using the governing equations of motion for the system. The equations are derived in [9]. In this
+paper, the cart pole problem was solved using both MPC control with the equations of motion and using MPC control in MuJoCo in order to compare the results.
+The states for each cart pole problem are shown in \ref{eq:cartpolestates} and the initial and target states of this problem are shown in
+\ref{tab:init_tar_carteqn}.
+
+\begin{equation}
+  \mb{x} = 
+  \begin{bmatrix}
+  \\x\\
+  \\\theta\\
+  \\\dot{x}\\
+  \\\dot{\theta}\\
+  \end{bmatrix}, \quad
+  \label{eq:cartpolestates}
+\end{equation}
+
+\begin{table}[h]
+  \centering
+  \caption{Initial and Target States of the Cart Pole Problem.}
+  \begin{tabular}{lrr}
+  \toprule
+  State & Initial & Target\\
+  \midrule
+  $x$ & 0  & 0\\
+  $\theta$ & 0 & $\pi$ \\
+  $\dot{x}$ & 0 & 0\\
+  $\dot{\theta}$ & 0 & 0\\
+  \end{tabular}
+  \label{tab:init_tar_carteqn}
+\end{table}
+
+Both methods of solving the cart pole problem used a similar optimization problem as shown in \ref{eq:cartpoleopt}. 
+Solving the cart pole problem using the equations of motion was straight forward as the equations of motion were well known and could be easily implemented in do-mpc.
+However, solving the cart pole problem in MuJoCo required a different approach in order to obtain the equations of motion for the states as the equations of motion 
+were assumed to be unknown as they would be in the case of a humanoid robot or other complex system.
+
+\begin{equation}
+  \begin{aligned}
+  &\text{Objective: $arg min_{x_k,u_k} (x_N - x_r)^TQ_N(x_N - x_r) + \sum_{k = 0}^{N-1} (x_k - x_r)^TQ(x_k - x_r) + u_k^TRu_k $}\\
+  &\text{Subject to:}\\
+  &\text{Control Range: $-35 < F < 35$}\\
+  &\text{Equations of motion for respective method}\\
+  &\text{State Constraints: $-3.5 < x < 3.5$}
+  \end{aligned}
+  \label{eq:cartpoleopt}
+\end{equation}
+
+In order to obtain the equations of motion for the MuJoCo cart pole system, a finite difference approach was used in order to solve for the linearized equations of motion at 
+each time step in the simulation. These equations took the form of \ref{eq:mj_eqn_form}, where the A and B matrices were equal to the jacobian of the system 
+with respect to the states and 
+control, respectively.
+
+\begin{equation}
+  \dot{x} = Ax + Bu
+  \label{eq:mj_eqn_form}
+\end{equation}
+
+The A and B matrices were found using the finite difference method by first obtaining the nominal (current, found from MuJoCo simulation)
+state of the system as denoted in \ref{eq:nomstate}
+and then perturbing each state one at a time as shown in \ref{eq:pertA} (repeat for all columns of the square matrix A) where $\epsilon$
+is a small number in order to obtain the A matrix. The B matrix was 
+obtained in a similar manner by perturbing the control input as shown in \ref{eq:pertB}.
+
+\begin{equation}
+  x_{nom} = f(x^0, \theta^0, \dot{x}^0, \dot{\theta}^0, F^0)
+  \label{eq:nomstate}
+\end{equation}
+
+\begin{equation}
+  A \text{ (Column One)} =\frac{ f(x^0 + \epsilon, \theta^0, \dot{x}^0, \dot{\theta}^0, F^0) - f(x^0, \theta^0, \dot{x}^0, \dot{\theta}^0, F^0)}{\epsilon}
+  \label{eq:pertA}
+\end{equation}
+
+\begin{equation}
+  B =\frac{ f(x^0, \theta^0, \dot{x}^0, \dot{\theta}^0, F^0 + \epsilon) - f(x^0, \theta^0, \dot{x}^0, \dot{\theta}^0, F^0)}{\epsilon}
+  \label{eq:pertB}
+\end{equation}
+
+After determining the A and B matrices for the cart pole MuJoCo system, the problem could then be solved by integrating the MuJoCo engine with the do-mpc interface.
+However, since the A and B matrices were being calculated at each time step, the dompc problem needed to be re-initialzed at each time step with the updated A and B matrices.
+This caused computational inefficency and will be addressed in future research.
+
+\section{Results}
+
+\subsection{Phase I: Compass Gait and Three Legged Walker}
 

### 2025-01-07 article_fall24/main.tex
paper update

@@
 \usepackage{appendix}
+
 % \usepackage{geometry}
@@
 
-This portion of research can be broken down into three phases; phase one involved using the work done in [6] and implementing MPC control on both the two legged walker 
-as well as a slightly more complicated three legged walker. The purpose of this phase was to understand how to implement MPC control on a simplified model of a robotic system.
+This portion of research can be broken down into three phases; phase one involved using the work done in [6] and implementing MPC control on a slightly 
+more complicated three link walker. The purpose of this phase was to understand how to implement MPC control on a simplified model of a robotic system.
 Phase two of this research consisted of using the knowledge of MPC control from phase one and implementing it in a more complex system using the MuJoCo physics engine.
@@
 
-\subsection{Phase I: Compass Gait and Three Legged Walker}
+\subsection{Phase I: MPC of the Three Link Walker}
 
-Phase I consisted of building on the research done in [6] and implementing MPC control on the compass gait system as well as a slightly more complex three legged walker.
-The compass gait is a simple two legged walking robot which is able to passively walk down an incline with no control input. The three legged walker is a similar system 
-but is more realistic as it uses a knee joint to walk. The compass gait as well as three legged walker are hybrid systems, meaning they change dynamics based on the state of the system.
+Phase I consisted of building on the research done in [6] and implementing MPC control a slightly more complex three link walker.
+The compass gait is a simple two legged walking robot which is able to passively walk down an incline with no control input. The three link walker is a similar system 
+but is more realistic as it uses a knee joint to walk. The compass gait as well as three link walker are hybrid systems, meaning they change dynamics based on the state of the system.
 The compass gait is shown in Figure \ref{fig:compass_gait}; for a more detailed look at the dynamics of the compass gait, see [6].
@@
 
-The three legged walker system is shown in Figure \ref{fig:threeleg}. For a more detailed look at the dynamics of the three legged walker, see [8]. In both the three legged walker and
+The three link walker system is shown in Figure \ref{fig:threeleg}. For a more detailed look at the dynamics of the three link walker, see [8]. In both the three link walker and
 compass gait, there is a single control input applied to the hip.
-
-In order to implement MPC control on the compass gait, do-mpc was used with the optimization problem as shown in \ref{eq:twolegopt} where the states of the system ($\mb{x}$)
-are shown in \ref{eq:twolegstates}. In the objective function $x_N$ and $x_k$ represent the terminal and the current state of the system, respectively.
-$x_r$ represents the target state of the system, which is determined to be the state at which both the swing and stance legs are in contact with the ground. This state 
-is shown in Table \ref{tab:twolegterm}.
-
-\begin{equation}
-  \begin{aligned}
-  &\text{Objective: $arg min_{x_k,u_k} (x_N - x_r)^TQ_N(x_N - x_r) + \sum_{k = 0}^{N-1} (x_k - x_r)^TQ(x_k - x_r) + u_k^TRu_k $}\\
-  &\text{Subject to:}\\
-  &\text{Dynamic Equations of System}\\
-  &\text{State Constraints: $-90^{\circ} < x_1 < 90^{\circ }, -90^{\circ} < x_2 < 90^{\circ }  $}
-  \end{aligned}
-  \label{eq:twolegopt}
-\end{equation}
-
-\begin{equation}
-  \mb{x} = 
-  \begin{bmatrix}
-  \\x_{sw}\\
-  \\x_{st}\\
-  \\\dot{x}_{sw}\\
-  \\\dot{x}_{st}\\
-  \end{bmatrix}, \quad
-  \label{eq:twolegstates}
-\end{equation}
-
-\begin{table}[h]
-  \centering
-  \caption{Target states during MPC control of the compass gait.}
-  \begin{tabular}{lr}
-  \toprule
-  State & Target Value \\
-  \midrule
-  $x_{sw}$ & 0.19 \\
-  $x_{st}$ & -0.3 \\
-  \end{tabular}
-  \label{tab:twolegterm}
-\end{table}
 
@@
   \includegraphics[width=8cm]{./figures/threeleg.png}
-  \caption{Diagram of the three legged walker model where q1, q2, and q3 are the stance, swing, and knee joint angles.}
+  \caption{Diagram of the three link walker model where q1, q2, and q3 are the stance, swing, and knee joint angles.}
   \label{fig:threeleg}
@@
 
-The $Q_N$ and $Q$ matrices are the weighting matrices for the MPC control and are simply set to be equal to the identity matrix in this case. 
-The $R$ matrix is the weighting matrix for the control input and is set to be equal to 10 to penalize the single control input.
-
-MPC control was implemented on the three legged walker in a similar manner to the compass gait, except there were two phases considered for this system, a two link and a 
-three link phase. A diagram showing these phases is shown in Figure \ref{fig:threelegstates}. These phases were treated as seperate systems with different governing equations and 
+MPC control was implemented on the three link walker by considering two phases for this system, a two link and a 
+three link phase, where the two link phase has the same dynamics as the compass gait. A diagram showing these phases is shown in 
+Figure \ref{fig:threelegstates}. These phases were treated as seperate systems with different governing equations and 
 each phase was controlled using a seperate MPC controller. The transition between the two phases was triggered by "knee-strike", where the 
@@
   \includegraphics[width=8cm]{./figures/threelegstates.png}
-  \caption{Diagram showing the two link and three link phases of the three legged walker.}
+  \caption{Diagram showing the two link and three link phases of the three link walker.}
   \label{fig:threelegstates}
@@
 
-Similarly to the compass gait control problem, the $Q_N$ and $Q$ matrices are the identity matrix and the $R$ matrix is set to 1 in this case.
+In the optimization problem, $x_N$ and $x_k$ represent the terminal and current states of the system and $x_r$ respresents the target states.
+$N$ is defined as the finite horizon of the MPC problem as can be thought of as how many timesteps in the future the model predicts the behavior of the system.
+the $Q_N$ and $Q$ matrices are weighting matrices for the control of each state; In this case, they are equal to the identity matrix.
+The $R$ matrix is a weighting matrix to penalize the control input and is equal to 1 (scalar for one control input).
 The target states for each phase are the states which were pre-determined to be the knee strike and heel strike states and are shown in Table \ref{tab:threelegterm}.
@@
   \centering
-  \caption{Target states during MPC control of the three legged walker.}
+  \caption{Target states during MPC control of the three link walker.}
   \begin{tabular}{lrr}
@@
 Phase II of this research consisted of using the knowledge of MPC which was gained from phase I and applying it to a more complex system in the MuJoCo physics engine which can 
-simulate complicated robotic systems such as the humanoid shown in \ref{fig:humanoid}. Instead of applying MPC control to the humanoid, MPC control was first applied to a simpler system 
+simulate complicated robotic systems such as the humanoid shown in Figure \ref{fig:humanoid}. Instead of applying MPC control to the humanoid, MPC control was first applied to a simpler system 
 in a way which can be easily scaled up to the humanoid. The chosen system was the classic cart pole control problem, where a pole with a 
-mass attatched to the end is attatched to a cart which can move back and forth in one direction. This is shown in \ref{fig:cartpole}. The goal of this problem is to 
+mass attatched to the end is attatched to a cart which can move back and forth in one direction. This is shown in Figure \ref{fig:cartpole}. The goal of this problem is to 
 control the pole to balance upright on the cart when the cart is at a stationary set location. This is done by applying a force to the cart to move it in the positive or 
@@
 
-\subsection{Phase I: Compass Gait and Three Legged Walker}
+\subsection{Phase I: MPC of the Three Link Walker}
+
+The three link walker was simulated using the MPC control scheme outlined in the previous section. The 
+walker was also simulated without control in order to compare the limit cycles of the simulations. The parameters used for simulation are shown in Table
+\ref{tab:walker_params} and the initial condition is shown in Table \ref{tab:walker_init}
+The gait of the passive walker as well as 
+the timeseries states are shown in Figure \ref{fig:passive_sol}. As can be seen, both the knee strike and heel strike can be seen clearly in the gait cycle.
+The knee strike occurs as a small instantanious change in angular velocity of the swing leg (the $x_2, x_5$ leg). The heel strike occurs when the system switches the swing legs 
+and the stance legs. This can be seem as the transition from the $x_2, x_5$ graph to the $x_1, x_4$ graph.
+
+\begin{table}[h]
+  \centering
+  \caption{Parameters of the Three Link Walker Used in Simulation.}
+  \begin{tabular}{lr}
+  \toprule
+  Parameter & Knee Strike Target\\
+  \midrule
+  $a_1$ & 0.375\\
+  $b_1$ & 0.125\\
+  $a_2$ & 0.175\\
+  $b_2$ & 0.325\\
+  $m_h$ & 0.5\\
+  $m_t$ & 0.5\\
+  $m_s$ & 0.05\\
+  $\gamma$ & 0.05\\
+  \end{tabular}
+  \label{tab:walker_params}
+\end{table}
+
+\begin{table}[h]
+  \centering
+  \caption{Initial Condition of the three link walker.}
+  \begin{tabular}{lrr}
+  \toprule
+  State & Initial Condition\\
+  \midrule
+  $x_{1}$ & 0.1877 \\
+  $x_{2}$ & -0.2884\\
+  $x_{3}$ & -0.2884\\
+  $x_{4}$ & -1.1014\\
+  $x_{5}$ & -0.0399\\
+  $x_{6}$ & -0.0399\\
+  \end{tabular}
+  \label{tab:walker_init}
+\end{table}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=.4\linewidth]{./figures/gait_passive.png} % Just stack two includegraphics!
+  \includegraphics[width=.4\linewidth]{./figures/ts_passive.png}
+  \caption{The Gait cycle and timeseries states of the passive three link walker}
+  \label{fig:passive_sol}
+  \end{figure}
+
+The MPC controlled graphs for the three link walker are similarly shown in Figure \ref{fig:active_sol}. The chaotic nature of the gait cycle in the active three link walker 
+can be clearly seen. Despite the walker acheiving the goal of not falling over, there are clear stability issues with the MPC controller. This is most likely due 
+to the way the objective of control is defined. Since the target states are simply defined as predetermined states where knee strike and heel strike occur, the 
+active walker is constrained in its motion and is unable to find a stable gait cycle which passes through the specified transition points. 
+Using a different control objective such as controlling the average forward speed of the walker or doing path planning to find the optimal path the legs should take 
+would most likely cause the gait cycle of this walker to be stable.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=.4\linewidth]{./figures/three_gait.png} % Just stack two includegraphics!
+  \includegraphics[width=.4\linewidth]{./figures/timeseries_data.png}
+  \caption{The Gait cycle and timeseries states of the active three link walker}
+  \label{fig:active_sol}
+\end{figure}
+
+\subsection{MuJoCo MPC}
+The cartpole system was simulated using MPC control for both cases, using the equations of motion and MuJoCo. The parameters for the cartpole using the equations of motion are shown 
+in Table \ref{tab:cartpole_params}. The MuJoCo case used the same pole length and cart mass; however, it should be noted that the physics of the system were slightly different.
+In the MuJoCo system, the pole was treated as a solid rod of constant density, while in the equations of motion case, the pole was treated as a massless rod 
+with a point mass attatched to the top of it. This changes the dynamics of the system and should be addressed in future research.
+
+\begin{table}[h]
+  \centering
+  \caption{Parameters Used in the Cart Pole Simulation (Eqs of Motion)}
+  \begin{tabular}{lrr}
+  \toprule
+  Parameter & Value\\
+  \midrule
+  Cart mass & 0.6 \\
+  Pole mass & 0.2\\
+  Pole Length & 0.5\\
+  \end{tabular}
+  \label{tab:cartpole_params}
+\end{table}
+
+The timeseries solution of the equations of motion case is shown in Figure \ref{fig:sol_eqn}. As can be seen, the system reaches a steady state with the pole balanced 
+at the upright position and the cart at the target state. The timeseries solution of the MuJoCo case is shown in Figure \ref{fig:sol_mj}. In the MuJoCo case, it can 
+also be seen that the pole is balanced in the upright position at the end of the simulation with the cart in the target state (in the MuJoCo system, the 
+$\theta$ axis is offset by $\pi$, meaning the target and initial states are switched). It can also be seen that the MuJoCo controller uses more total force to 
+balance the cartpole and takes a longer time to reach steady state. This is most likely due to the difference in physics between the two systems and will be addressed in the future.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=8cm]{./figures/timeseries_dynamics.png} % Just stack two includegraphics!
+  \caption{The timeseries solution to the cartpole problem (using equations of motion)}
+  \label{fig:sol_eqn}
+\end{figure}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=8cm]{./figures/cartpole_mj_timeseries_normlen.png} % Just stack two includegraphics!
+  \caption{The timeseries solution to the cartpole using MuJoCo}
+  \label{fig:sol_mj}
+\end{figure}
+
+In Figure \ref{fig:sol_mj} the calculated states from the MuJoCo simulation and the states from the linearized equations of motion (using 
+finite difference) are compared. As can be seen the MuJoCo states are equal to the linearized equations of motion states (do-mpc states) throughout the duration 
+of the simulation. One thing that should be considered however, is that when using MPC control, the future states of the system are predicted for a set 
+"control horizon" (70 time steps in this case). Since the system is linearized at each time step during the simulation, it is expected that the current state of the 
+system is equivelant between the MuJoCo simulation and the linearized equations of motion; however, during the control horizon it is probable that after a certain amount of time 
+when the state of the system is sufficently far from the state about which the system was linearized that the predictions become inaccurate. This means that it is possible 
+for a control horizon which is too long to cause the control to fail due to the inaccuracy in the future predictions of the model.
+
+While the MPC controller works to stabilize the MuJoCo system when the parameters are the same as the above case, it was observed 
+that if the parameters are changed too far from the original parameters the control will fail. An example of this using the parameters shown in 
+\ref{tab:fail_cartpole} is shown in Figure \ref{fig:fail_cartpole}. As can be seen, the pole position is unstable and does not balance. It is hypothesied that 
+this failure is caused by the difference in physics between the two cases, as the equations of motion case does not fail when the parameters are changed. It is also possible 
+that the inaccurate control horizon is causing this failure. This is being investigated by creating a new MuJoCo model which is exactly the same as the equations of motion case 
+and comparing the two cases again.
 

### 2025-01-07 code/nl_mpc_example/dompc/cartpole_dynamics/main.py
update images

@@ -12,11 +12,11 @@ import do_mpc
 from model import model_set
 from controller import control
 
-m = 0.2
-M = 0.6
-L = 0.5
+m = 5
+M = 10.5
+L = 0.3
 
-num_steps = 200
+num_steps = 400
 
 delta_t = .04
 model = model_set(M,m,L)

### 2025-01-07 code/nl_mpc_example/dompc/cartpole_mujoco/main.py

@@ -32,6 +32,7 @@ def balance(delta_t = 0.02, plotting = False, polelen = 0.3, kg_m_pole = 16.687)
  
     # set pole length and update mass/intertia
     model.body_mass[2] = polelen*kg_m_pole
+    print(model.body_mass)
     model.body_inertia[1,2] = (1/12)*polelen*kg_m_pole*(polelen**2)
     #model.body_mass[1] = 9
 
@@ -187,4 +188,4 @@ def balance(delta_t = 0.02, plotting = False, polelen = 0.3, kg_m_pole = 16.687)
 
 
 if __name__ == "__main__":
-    balance(polelen = 0.3,plotting=True)
+    balance(polelen = 0.2,plotting=True)

### 2025-01-07 article_fall24/main.tex
complete article

@@
 While the MPC controller works to stabilize the MuJoCo system when the parameters are the same as the above case, it was observed 
-that if the parameters are changed too far from the original parameters the control will fail. An example of this using the parameters shown in 
-\ref{tab:fail_cartpole} is shown in Figure \ref{fig:fail_cartpole}. As can be seen, the pole position is unstable and does not balance. It is hypothesied that 
+that if the parameters are changed too far from the original parameters the control will fail. Shown are two examples of this in 
+\ref{fig:cartpole_fail} where the pole length is changed to 0.5 and 0.2m.
+As can be seen, the pole position is unstable and does not balance. It is hypothesied that 
 this failure is caused by the difference in physics between the two cases, as the equations of motion case does not fail when the parameters are changed. It is also possible 
@@
 
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=6cm]{./figures/cartpole_mjpc_times_0.5.png} % Just stack two includegraphics!
+  \includegraphics[width=6cm]{./figures/cartpole_mjpc_times_0.2.png} % Just stack two includegraphics!
+  \caption{Control of the MuJoCo cartpole fails with pole lengths of 0.5m (left) and 0.2m (right).}
+  \label{fig:cartpole_fail}
+\end{figure}
+
+\section{Conclusion and Future Work}
+The two phases of research discussed in this paper include implementing MPC control on a three link walker in order to 
+better understand how MPC control can be applied to simple robotic systems using do-mpc as well as 
+implementing MPC control on a general system (the cartpole problem) in the MuJoCo physics engine in order to eventually scale MPC control to a more 
+complex robotic system in MuJoCo such as the humanoid robot in phase III of this reserach.
+
+MPC control was implemented on a three link walker using do-mpc with the goal of steadily walking down a slope. This goal was acheived by the walker; however,
+the gait cycle of the walker was chaotic and unstable. This is most likely due to a poor choice of control objective where the target states were simply predetermined 
+states where knee strike and heel strike occur. In future work such as implementing MPC control on the humanoid robot, a better control objective will be established such as 
+controlling the average speed of the humanoid and also implementing path planning and motion planning in order to acheive a stable gait cycle.
+
+MPC control was also established over the cartpole system in MuJoCo in phase II of this research. The MuJoCo cartpole system solution was compared to a solution
+where the equations of motion of the cartpole were used to simulate the system instead of a physics engine. In both cases, the cartpole was able to be balanced when the 
+parameters of the system were ideal. However, in the MuJoCo simulation case, the control of the cartpole was not robust and failed when the parameters of the system were
+poorly chosen while the equations of motion case did not fail under the same conditions. It is hypothesied that this is due to the difference in physics 
+between the two systems as the MuJoCo system treats the pole as a solid rod of constant density while the equations of motion case treats the pole as a massless rod with a point mass 
+attatched to the end. This change in dynamics causes a drastic difference in the moment of inertia of the pole and makes it much more difficult to control in the 
+MuJoCo case. This hypothesis is being tested in future research by creating a new MuJoCo model which is exactly the same as the equations of motion case 
+and comparing the two cases again.
+
+Phase III of this research involves scaling the MPC control of the cartpole system in MuJoCo up into a more complex humanoid robot system and preforming CCD on the robot. 
+This will involve implementing better algorithims for control such as path and motion planning and obstacle avoidance. This will also involve implementing a different 
+MPC controller rather than using do-mpc in order to eliminate the inefficencies caused by re-initializing the do-mpc system at each time step in the simulation. The end 
+goal of this research is to create a tool which can be used to preform robust control co-design on any dynamical system (linear or non-linear) using MPC control.
+
 \end{document}

### 2025-01-07 article_fall24/main.tex
paper again

@@
 Control Co-Design (CCD) is an engineering design strategy used to optimize the design of a system with control in mind. CCD can be acheived using various control schemes, 
-including open loop control [1], PID control, and other optimal control methods such as LQR or MPC Control. CCD has a very wide range of applications such as in 
-aerospace systems [2] and in thermal energy storage [3]. While CCD has been used in various industries there has yet to be a compreherensive tool developed
+including open loop control \cite{Sundarrajan2023}, PID control, and other optimal control methods such as LQR or MPC Control. CCD has a very wide range of applications such as in 
+aerospace systems \cite{Van2019} and in thermal energy storage \cite{10177617}. While CCD has been used in various industries there has yet to be a compreherensive tool developed
 to preform robust CCD using differentiable optimal control strategies such as nonlinear MPC control.
@@
 This research focuses on developing a tool which can be used in order to preform robust CCD using MPC control. This research is primarly focused on robotic applications and uses 
-the MuJoCo [4] physics simulation engine in order to simulate the robotic systems. In this paper, do-mpc [5], a library for nonlinear MPC control, is used to simplify the 
-implementation of MPC control. This research builds off of previous work [6] which used the Dymos Optimal Control library [7] in order to preform CCD of a simple 
+the MuJoCo \cite{todorov2012mujoco} physics simulation engine in order to simulate the robotic systems.
+In this paper, do-mpc \cite{FIEDLER2023105676}, a library for nonlinear MPC control, is used to simplify the
+implementation of MPC control. This research builds off of previous work \cite{Howell2024} which used the Dymos Optimal Control library \cite{Falck2021} in order to preform CCD of a simple 
 two legged walking robot known as the compass gait.
 
-This portion of research can be broken down into three phases; phase one involved using the work done in [6] and implementing MPC control on a slightly 
+This portion of research can be broken down into three phases; phase one involved using the work done in \cite{Howell2024} and implementing MPC control on a slightly 
 more complicated three link walker. The purpose of this phase was to understand how to implement MPC control on a simplified model of a robotic system.
@@
 
-Phase I consisted of building on the research done in [6] and implementing MPC control a slightly more complex three link walker.
+Phase I consisted of building on the research done in \cite{Howell2024} and implementing MPC control a slightly more complex three link walker.
 The compass gait is a simple two legged walking robot which is able to passively walk down an incline with no control input. The three link walker is a similar system 
 but is more realistic as it uses a knee joint to walk. The compass gait as well as three link walker are hybrid systems, meaning they change dynamics based on the state of the system.
-The compass gait is shown in Figure \ref{fig:compass_gait}; for a more detailed look at the dynamics of the compass gait, see [6].
+The compass gait is shown in Figure \ref{fig:compass_gait}; for a more detailed look at the dynamics of the compass gait, see \cite{Howell2024}.
 
@@
 
-The three link walker system is shown in Figure \ref{fig:threeleg}. For a more detailed look at the dynamics of the three link walker, see [8]. In both the three link walker and
+The three link walker system is shown in Figure \ref{fig:threeleg}. For a more detailed look at the dynamics of the three link walker, see \cite{Chen2007PassiveDW}. In both the three link walker and
 compass gait, there is a single control input applied to the hip.
@@
 knee joint angle became equal to the angle of the swing leg, and "heel-strike", where the swing leg comes into contact with the ground. The 
-transition between the phases were calculated using the transition equations from [8].
+transition between the phases were calculated using the transition equations from \cite{Chen2007PassiveDW}.
 \begin{figure}[!h]
@@
 
-The cart pole system has previously been solved with various control techniques using the governing equations of motion for the system. The equations are derived in [9]. In this
+The cart pole system has previously been solved with various control techniques using the governing equations of motion for the system. The equations are derived in \cite{key}. In this
 paper, the cart pole problem was solved using both MPC control with the equations of motion and using MPC control in MuJoCo in order to compare the results.
@@
 
+\pagestyle{plain}
+\normalsize
+\bibliography{bib/references}
+
 \end{document}

### 2025-01-08 code/gym/PID_cartpole.py
double pendulum mpc

@@ -1,59 +1,0 @@
-import mujoco
-import glfw
-import numpy as np
-np.set_printoptions(precision=4)
-import matplotlib.pyplot as plt
-
-def init_window(max_width, max_height):
-    glfw.init()
-    window = glfw.create_window(width=max_width, height=max_height,
-                                       title='Demo', monitor=None,
-                                       share=None)
-    glfw.make_context_current(window)
-    return window
-
-window = init_window(2400, 1800)
-width, height = glfw.get_framebuffer_size(window)
-viewport = mujoco.MjrRect(0, 0, width, height)
-
-xml = "/home/max/workspace/research_template/code/gym/inverted_pendulum.xml"
-
-model = mujoco.MjModel.from_xml_path(xml)
-data = mujoco.MjData(model)
-context = mujoco.MjrContext(model, mujoco.mjtFontScale.mjFONTSCALE_100)
-
-scene = mujoco.MjvScene(model, 6000)
-camera = mujoco.MjvCamera()
-camera.trackbodyid = 2
-camera.distance = 3
-camera.azimuth = 90
-camera.elevation = -20
-mujoco.mjv_updateScene(
-    model, data, mujoco.MjvOption(), mujoco.MjvPerturb(),
-    camera, mujoco.mjtCatBit.mjCAT_ALL, scene)
-
-from pid_control import PD,Controller
-controller = Controller()
-
-data.qpos = np.array([-0.1, np.deg2rad(5)])
-
-
-
-while(not glfw.window_should_close(window)):
-    mujoco.mj_step1(model, data)
-
-    data.ctrl = -controller.observe(data.qpos[0],data.qpos[1])
-    
-    print(data.ctrl)
-    mujoco.mj_step2(model, data)
-
-    mujoco.mjv_updateScene(
-        model, data, mujoco.MjvOption(), None,
-        camera, mujoco.mjtCatBit.mjCAT_ALL, scene)
-    mujoco.mjr_render(viewport, scene, context)
-
-    glfw.swap_buffers(window)
-    glfw.poll_events()
-
-glfw.terminate()
-

### 2025-01-08 code/gym/mjpc_tests/particle.py

@@ -1,131 +1,0 @@
-# Copyright 2023 DeepMind Technologies Limited
-#
-# Licensed under the Apache License, Version 2.0 (the "License");
-# you may not use this file except in compliance with the License.
-# You may obtain a copy of the License at
-#
-#     http://www.apache.org/licenses/LICENSE-2.0
-#
-# Unless required by applicable law or agreed to in writing, software
-# distributed under the License is distributed on an "AS IS" BASIS,
-# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-# See the License for the specific language governing permissions and
-# limitations under the License.
-# ==============================================================================
-
-import mediapy as media
-import mujoco
-import predictive_sampling
-import numpy as np
-
-
-# %%
-xml = '/home/max/workspace/research_template/code/gym/mjpc_tests/xmls/particle.xml'
-
-# create simulation model + data
-model = mujoco.MjModel.from_xml_path(xml)
-data = mujoco.MjData(model)
-#renderer = mujoco.Renderer(model)
-
-
-# %%
-# reward
-def reward(model: mujoco.MjModel, data: mujoco.MjData) -> float:
-  # position
-  goal = data.mocap_pos[0, :2]
-  pos_error = data.qpos - goal
-  r0 = -np.dot(pos_error, pos_error)
-
-  # velocity
-  r1 = -np.dot(data.qvel, data.qvel)
-
-  # effort
-  r2 = -np.dot(data.ctrl, data.ctrl)
-
-  return 5.0 * r0 + 0.1 * r1 + 0.1 * r2
-
-
-# %%
-# planner
-horizon = 0.5
-splinestep = 0.1
-planstep = 0.025
-nimprove = 4
-nsample = 4
-noise_scale = 0.01
-interp = "zero"
-planner = predictive_sampling.Planner(
-    model,
-    reward,
-    horizon,
-    splinestep,
-    planstep,
-    nsample,
-    noise_scale,
-    nimprove,
-    interp=interp,
-)
-# %%
-# simulate
-mujoco.mj_resetData(model, data)
-steps = 301
-
-# set goal position
-data.mocap_pos[0, :2] = np.array([0.3, 0.0])
-
-# history
-qpos = [data.qpos]
-qvel = [data.qvel]
-act = [data.act]
-ctrl = []
-rewards = []
-
-# frames
-frames = []
[…]

### 2025-01-08 code/gym/pid_control.py

@@ -1,28 +1,0 @@
-class PD:
-    def __init__(self, kp, kd, goal):
-        self.kp = kp
-        self.kd = kd
-        self.goal = goal
-        self.last_error = 0
-
-    def observe(self, x):
-        error = self.goal - x
-        d_error = error - self.last_error
-        self.last_error = error
-        return self.kp * error + self.kd * d_error
-    
-class Controller:
-    def __init__(self):
-        self.cart = PD(kp=1, kd=100, goal=0)
-        self.pole = PD(kp=5, kd=100, goal=0)
-
-    def observe(self, cart_position, pole_angle):
-        u_cart = self.cart.observe(cart_position)
-        u_pole = self.pole.observe(pole_angle)
-        if u_pole+u_cart > 0:
-            action = 1
-        elif u_pole+u_cart <0:
-            action = -1
-        else:
-            action = 0
-        return action

### 2025-01-08 code/nl_mpc_example/dompc/cartpole_mujoco/main.py

@@ -24,7 +24,7 @@ from OpenGL.GLU import *
 from OpenGL.GLUT import *
 
 # make function to do co design using finite difference
-def balance(delta_t = 0.02, plotting = False, polelen = 0.3, kg_m_pole = 16.687):
+def balance(delta_t = 0.01, plotting = False, polelen = 0.3, kg_m_pole = 16.687):
 
     # set initial conditions
     x0 = [0, np.deg2rad(180)]
@@ -188,4 +188,4 @@ def balance(delta_t = 0.02, plotting = False, polelen = 0.3, kg_m_pole = 16.687)
 
 
 if __name__ == "__main__":
-    balance(polelen = 0.2,plotting=True)
+    balance(polelen = 0.3,plotting=True)

### 2025-01-08 code/nl_mpc_example/dompc/doublepend_mujoco/controller.py

@@ -14,7 +14,7 @@ def control (model, delta_t):
     mpc = do_mpc.controller.MPC(model)
 
     setup_mpc = {
-        'n_horizon': 400,
+        'n_horizon': 50,
         'n_robust': 0,
         'open_loop': 0,
         't_step': delta_t,
@@ -24,7 +24,7 @@ def control (model, delta_t):
         'collocation_ni': 1,
         'store_full_solution': True,
         # Use MA27 linear solver in ipopt for faster calculations:
-        'nlpsol_opts': {'ipopt.linear_solver': 'mumps'}
+        # 'nlpsol_opts': {'ipopt.linear_solver': 'mumps'}
     }
     mpc.settings.supress_ipopt_output()
     mpc.set_param(**setup_mpc)
@@ -37,8 +37,8 @@ def control (model, delta_t):
 
     mpc.set_rterm(u=0.001) # input penalty
 
-    max_x = np.array([[3.5], [200], [1000], [1000]])
-    min_x = np.array([[-3.5], [-200], [-1000], [-1000]])
+    max_x = np.array([[1000], [1000], [1000], [1000]])
+    min_x = np.array([[-1000], [-1000], [-1000], [-1000]])
 
     # lower bounds of the states
     mpc.bounds['lower','_x','x'] = min_x
@@ -47,10 +47,10 @@ def control (model, delta_t):
     mpc.bounds['upper','_x','x'] = max_x
 
     # lower bounds of the input
-    mpc.bounds['lower','_u','u'] = -35
+    mpc.bounds['lower','_u','u'] = -10
 
     # upper bounds of the input
-    mpc.bounds['upper','_u','u'] =  35
+    mpc.bounds['upper','_u','u'] =  10
 
     mpc.setup()
 
